### 2023-11-22 Scripts/Data_Management/Scripts_Processing/Prep_TyrerCuzick_input.Rmd
Initial commit

---
title: "Prepare data inputs for Tyrer-Cuzick software v8"
author: "Jennifer Collister"
date: "`r Sys.Date()`"
output: 
  word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)
library(ggplot2)
library(tidyr)
library(readr)
library(dplyr)
library(glue)

# Pretty plots
library(etm)

# Pretty tables
library(arsenal)
library(pander)

# Survival 
library(survival)
library(survminer)

# Flowchart
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)
library(png)

# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(file.path(config$functions)))

#-------------------------------------------------------------------------------------------------------------------

# Output Formatting

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)

# Pander options
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')

# Set formatting options for tableby objects
my_controls <- tableby.control(
    test = F,
    total = T,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "Nmiss"), #"medianq1q3", 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      # medianq1q3 = "Median (IQR)",
      # range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

#-------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r derive-data, include=FALSE, cache=TRUE}

source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data) {

  excl$initial <<- nrow(data)

  # Restrict to females (by self-report and genetic)
  data <- data[data$BaC_Sex == "Female",]
  data <- data[data$GeP_Sex == "Female" & !is.na(data$GeP_Sex),]
  excl$sex <<- nrow(data)

  # Restrict to post-menopausal women
  data <- data[!is.na(data$FSF_MenopauseStatus) & data$FSF_MenopauseStatus == "Yes",]
  excl$meno <<- nrow(data)

  # Restrict to White British (genetically)
  # data <- data[!is.na(data$GeP_ethnic),]
  excl$whitebrit <<- nrow(data)

  # Exclude those outside the 40-70 age range
  data <- data[!is.na(data$TEU_BaC_AgeAtRec) & data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  excl$agerange <<- nrow(data)
  # Restrict population to those for whom we have PRS scores available
  data <- data[!is.na(data$PRS_bc_std),]
  excl$prs <<- nrow(data)
  # Exclude women with 20+ sisters as the TC model can't handle it
  data <- data[is.na(data$TC_NSisters) | data$TC_NSisters < 20,]
  # Exclude those who didn't answer HRT use
  data <- data[data$TC_HRT_raw != "Unanswered",]
  # Or who reported impossible HRT dates
  data <- data[is.na(data$FSF_HRTAgeStop) | data$FSF_HRTAgeStop <= ceiling(data$TEU_BaC_AgeAtRec),]
  excl$hrt <<- nrow(data)

  # Exclude prior breast cancer
  data <- data[is.na(data$BrCaDx_Prevalent),]
  data <- data[is.na(data$VI_BrCa),]
  excl$brca <<- nrow(data)
  # Prior mastectomy
  data <- data[data$TEU_HES_Mast_prev == "No",]
  excl$mast <<- nrow(data)
  # Prior carcinoma in situ
  data <- data[is.na(data$BrCaInSitu_Prevalent),]
  excl$cainsitu <<- nrow(data)

  return(data)
}

excl <- list(initial=0)

data <-
  evalWithMemoization(
    derive_variables(
      database = config$data$database,
      field_definitions = TEU_SPECS$TyrerCuzick,
      exclusions = exclusions
    ),
    key = c(TEU_SPECS$TyrerCuzick, exclusions)
  )

pretty_names <- pretty_switch(field_definitions=TEU_SPECS$TyrerCuzick)
pretty_list <- pretty_switch(field_definitions=TEU_SPECS$TyrerCuzick, return_type="list")

```

```{r}

table(data$TC_NSisters, useNA='ifany')

tcify <- function(data, nsis) {

  data <- data[data$TC_NSisters == nsis,]

  # Map data to correct format for Tyrer-Cuzick v8 model specification
  tc_data <- data %>% transmute(
    v1_ID = ID,
    v2_Age = floor(replace_na(TEU_BaC_AgeAtRec, -99)),
    v3_AgeMenarche = replace_na(FSF_AgeMenarche, -99),
    v4_Parity = dplyr::case_when(
      TC_Parity == "Nulliparous" ~ 0,
      TC_Parity == "Parous" ~ 1,
      TC_Parity == "Missing" ~ 2),
    v5_AgeFirstBirth = dplyr::case_when(
      TC_Parity == "Nulliparous" ~ 0,
      TC_Parity == "Missing" ~ 0,
      TC_Parity == "Parous" & !is.na(AgeFirstBirth) ~ as.numeric(AgeFirstBirth),
      TC_Parity == "Parous" & is.na(AgeFirstBirth) ~ -99
    ),
    v6_MenoStatus = 2, # we restricted to post-menopausal at baseline
    v7_MenoAge = replace_na(FSF_AgeMenopause, -99), # All women post-menopausal, so no need to handle unknown menopause status
    v8_Height = replace_na(Height, -99),
    v9_Weight = replace_na(Weight, -99),
    v10_Hyperplasia = 0, # missing or unknown
    v11_AtypHyperplasia = dplyr::case_when(
      AtypicalHyperplasia == "Yes" ~ 1,
      TRUE ~ 0
    ),
    v12_LCIS = 0, # missing or unknown - we excluded all CIS at baseline, very few were LCIS
    v13_NOTINUSE = 0, #
    v14_OvarianCancer = dplyr::case_when(
      TC_OvarianCaDx=="Yes" ~ 1,
      TC_OvarianCaDx=="No" ~ 0,
      ),
    v15_OvarianAge = floor(replace_na(TC_OvarianCaAge, -99)),
    v16_Ashkenazi = 0, # missing or unknown
    v17_HRTuse = dplyr::case_when(
      TC_HRT_imputed == "Never" ~ 0,
      TC_HRT_imputed == "Previous user (more than 5 years ago)" ~ 1,
      TC_HRT_imputed == "Previous user (less than 5 years ago)" ~ 2,
      TC_HRT_imputed == "Current" ~ 3
    ),
    v18_HRTType = 1, # missing or unknown
    v19_HRTDuration = round(replace_na(FSF_HRTDuration, 0),1),
    v20_HRTIntended = 0, # missing or unknown
    v21_HRTTimeSince = ifelse(!is.na(FSF_HRT_timesince) & FSF_HRT_timesince < 5,
                              round(FSF_HRT_timesince,1), 0),
    v22_GeneticTest = 0, # unknown

    # Family history
    v23_FatherGenetic = 0, # unknown
    v24_MotherBrCa = dplyr::case_when(
      TC_MotherBrCa=="Yes" ~ 1,
      TC_MotherBrCa=="No" ~ 0,
      ),
    v25_MotherBilateral = 0, # unknown
    v26_MotherOvarian = 0,
    v27_MotherBrCaAge = ifelse(TC_MotherBrCa == "Yes", -99, replace_na(TC_MotherAge, -99)),
    v28_MotherBilAge = -99,
    v29_MotherOvAge = -99,
    v30_MotherGenetic = 0, # unknown

    # Sisters
    v31_NSisters = replace_na(TC_NSisters, 0),
    # This gets complicated now, different number of columns depending on number of sisters
  )

  # This part is different depending on how many sisters
  sisters <- data %>% transmute(
    v32_SisterBrCa = dplyr::case_when(
      TC_SiblingBrCa=="Yes" ~ 1,
      TC_SiblingBrCa=="No" ~ 0,
      ),
    v33_SisterBilateral = 0,
    v34_SisterOvarian = 0,
    v35_SisterBrCaAge = -99,
    v36_SisterBilAge = -99,
    v37_SisterOvAge = -99,
    v38_SisterGenetic = 0
  )

  if(nsis == 1) {
    tc_data <- cbind(tc_data, sisters)
  } else if (nsis > 1) {
    extra <- sisters %>% mutate(v32_SisterBrCa = 0)
    for (n in c(1: (nsis-1))) {
      colnames(extra) <- str_replace(colnames(extra),
                                     "[0-9]+",
                                     function(x) {return(as.numeric(x) + 7)}
                                     )
      sisters <- cbind(sisters, extra)
    }
    tc_data <- cbind(tc_data, sisters)
  }

  # And the rest of the compulsory family history
  morefam <- data %>% transmute(
    # Paternal Gran
    s2_PatGranBrCa = 0,
    s3_PatGranBilateral = 0,
    s4_PatGranOvarian = 0,
    s5_PatGranBrCaAge = -99,
    s6_PatGranBilAge = -99,
    s7_PatGranOvAge = -99,
    s8_PatGranGenetic = 0,

    # Maternal gran
    s9_MatGranBrCa = 0,
    s10_MatGranBilateral = 0,
    s11_MatGranOvarian = 0,
    s12_MatGranBrCaAge = -99,
    s13_MatGranBilAge = -99,
    s14_MatGranOvAge = -99,
    s15_MatGranGenetic = 0,

    # Aunts
    s16_NPatAunts = 0,
    # s17_PatAuntBrCa = 0,
    # s18_PatAuntBilateral = 0,
    # s19_PatAuntOvarian = 0,
    # s20_PatAuntBrCaAge = -99,
    # s21_PatAuntBilAge = -99,
    # s22_PatAuntOvAge = -99,
    # s23_PatAuntGenetic = 0,

    # Maternal Aunts
    pa1_NMatAunts = 0,
    # pa2_MatAuntBrCa = 0,
    # pa3_MatAuntBilateral = 0,
    # pa4_MatAuntOvarian = 0,
    # pa5_MatAuntBrCaAge = -99,
    # pa6_MatAuntBilAge = -99,
    # pa7_MatAuntOvAge = -99,
    # pa8_MatAuntGenetic = 0,

    # Daughters
    ma1_NDaughters = 0,
    # ma2_DaughterBrCa = 0,
    # ma3_DaughterBilateral = 0,
    # ma4_DaughterOvarian = 0,
    # ma5_DaughterBrCaAge = -99,
    # ma6_DaughterBilAge = -99,
    # ma7_DaughterOvAge = -99,
    # ma8_DaughterGenetic = 0
  )
  tc_data <- cbind(tc_data, morefam)

  table3 <- data %>% transmute(
    zero1 = nsis,
    zero2 = 0,
    zero3 = 0,
    zero4 = 0,
    zero5 = 0,
    zero6 = 0,
    zero7 = 0
  ) # No participants have the optional family history data. Spec says 7 columns of 0
  if (nsis > 0) {
    for (sis in c(1:nsis)) {
      table3[paste0("sis", sis)] <- 0 
    }
  } # Technically these columns should come after zero1 but they're all just 0 so it doesn't matter
  tc_data <- cbind(tc_data, table3)

  v8 <- data %>% transmute(
    t1_FatherBrCa = dplyr::case_when(
      TC_FatherBrCa=="Yes" ~ 1,
      TC_FatherBrCa=="No" ~ 0,
      ), # All No but include anyway for completion
    t2_BrotherBrCa = dplyr::case_when(
      TC_SiblingBrCa == "Yes" & 
        (TC_NSisters == 0 | is.na(TC_NSisters)) & 
        (TC_NBrothers > 0 & !is.na(TC_NBrothers)) ~ 1,
      TRUE ~ 0
    ), # If siblings with breast cancer, no sisters, and brothers then yes, otherwise no
    t3_DensityMeasure = 0,
    t4_DensityValue = -99,
    t5_PRS = 0 #round(TEU_BrCa_313_PRS_center,8)
  )

  tc_data <- cbind(tc_data, v8)
  return(tc_data)
}

nrows <- nrow(data)
chunksize <- 15000
start <- 1
ordered <- data %>% # Treat NA sisters as 0
  mutate(TC_NSisters = ifelse(is.na(TC_NSisters), 0, TC_NSisters)) %>%
  arrange(TC_NSisters)

while(start < nrows) {

  end <- min(start + chunksize - 1, nrows)
  filename <- file.path(config$data$derived, "noPRS", paste0("rows", start, "_", end, ".txt"))

  chunk <- ordered[start:end,]
  nsisters <- sort(unique(chunk$TC_NSisters))
  nparticipants <- nrow(chunk)

  header <- data.frame(col1 = c("v8", nparticipants))
  write.table(header, file=filename, append=FALSE,
              sep="\t", eol="\n", quote=FALSE,
              row.names=FALSE, col.names=FALSE)

  for (n in nsisters) {
    output <- tcify(chunk, nsis=n)

    write.table(output, file=filename, append=TRUE,
              sep="\t", eol="\n", quote=FALSE,
              row.names=FALSE, col.names=FALSE)
  }
  start <- end + 1
}

#-------------#
# Run 3 times #
#-------------#
# t5_PRS = 0, filename "noPRS"
# t5_PRS = TEU_BrCa_313_PRS_center, filename "withPRS/Mavaddat2019"
# t5_PRS = TEU_BrCa_100k_PRS_center, filename = "withPRS/Fritsche2020"
# Note that the different PRS are on different scales

```

```{r}
# Read Tyrer-Cuzick risks output file

filenames <- list.files(file.path(config$data$derived, "noPRS", "risk"), 
                        full.names=TRUE, recursive=TRUE, include.dirs=FALSE)

allrisks <- lapply(filenames, function(x){
  read.table(x, header=FALSE, sep="\t",
             col.names=c("ID", "nyear", "ltime_risk", "pop_ltime_risk", "brca1", "brca2"))
  })

risks <- dplyr::bind_rows(allrisks)

```

### 2023-11-22 Scripts/Stats_Analysis/RMarkdown/Age_standardised_incidence.Rmd
Initial commit

---
title: "Variable Exploration"
author: "Jennifer Collister"
date: "`r Sys.Date()`"
output: 
  word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE, warning = FALSE, message = FALSE)

library(kableExtra)
library(knitr)
# library(yaml)
library(here)
library(ggplot2)
library(tidyr)
# library(readr)
library(dplyr)
# library(glue)
# library(BCRA)

# Pretty tables
# library(arsenal)
# library(pander)

# Survival 
# library(survival)
# library(survminer)
# library(rms)

# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
# config = yaml.load_file(here::here("./config.yml"))
# source(here::here(file.path(config$functions)))

#-------------------------------------------------------------------------------------------------------------------

# Output Formatting

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)

#-------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r derive-data, include=FALSE, cache=TRUE}

# source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data) {

  excl$initial <<- nrow(data)

  # Restrict to females (by self-report and genetic)
  data <- data[data$BaC_Sex == "Female",]
  data <- data[data$GeP_Sex == "Female" & !is.na(data$GeP_Sex),]
  excl$sex <<- nrow(data)

  # Restrict to post-menopausal women
  data <- data[!is.na(data$FSF_MenopauseStatus) & data$FSF_MenopauseStatus == "Yes",]
  excl$meno <<- nrow(data)

  # Restrict to White British (genetically)
  data <- data[!is.na(data$GeP_ethnic),]
  excl$whitebrit <<- nrow(data)

  # Exclude those outside the 40-70 age range
  data <- data[!is.na(data$TEU_BaC_AgeAtRec) & data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  excl$agerange <<- nrow(data)
  # Restrict population to those for whom we have PRS scores available
  data <- data[!is.na(data$PRS_bc_std),]
  excl$prs <<- nrow(data)
  # Exclude women with 20+ sisters as the TC model can't handle it
  data <- data[is.na(data$TC_NSisters) | data$TC_NSisters < 20,]
  # Exclude those who didn't answer HRT use
  data <- data[data$TC_HRT_raw != "Unanswered",]
  # Or who reported impossible HRT dates
  data <- data[is.na(data$FSF_HRTAgeStop) | data$FSF_HRTAgeStop <= ceiling(data$TEU_BaC_AgeAtRec),]
  excl$hrt <<- nrow(data)

  # Exclude prior breast cancer
  data <- data[is.na(data$BrCaDx_Prevalent),]
  data <- data[is.na(data$VI_BrCa),]
  excl$brca <<- nrow(data)
  # Prior mastectomy
  data <- data[data$TEU_HES_Mast_prev == "No",]
  excl$mast <<- nrow(data)
  # Prior carcinoma in situ
  data <- data[is.na(data$BrCaInSitu_Prevalent),]
  excl$cainsitu <<- nrow(data)

  return(data)
}

excl <- list(initial=0)

# data <- derive_variables(
#   database = config$data$database,
#   field_definitions = TEU_SPECS$TyrerCuzick,
#   exclusions = function(data){return(data[data$BaC_Sex == "Female" &
#                                             data$GeP_Sex == "Female" & !is.na(data$GeP_Sex),])}
#   )
# saveRDS(data, "K:/TEU/PRSforBrCa/Stats_Outputs/agestdinc.rds")
data <- readRDS("K:/TEU/PRSforBrCa/Stats_Outputs/agestdinc.rds")

wholepop <- data
analysis_pop <- exclusions(data)

```

# Age-specific and directly standardised breast cancer incidence rates

> We calculated age specific and directly standardised colorectal cancer incidence rates in UK Biobank overall and for the Integrated Modelling Cohort used to derive integrated QCancer-10+PRS risk models, and compared these with Office for National Statistics 2013 cancer registry data for England (chosen as the approximate mid-point of available UK Biobank follow-up).[15](https://www.ons.gov.uk/peoplepopulationandcommunity/healthandsocialcare/conditionsanddiseases/datasets/cancerregistrationstatisticscancerregistrationstatisticsengland) Age specific rates were calculated in five year age bands between 40 and 80 years as the number of first incident colorectal cancers over the number of person years at risk. Age standardised incidence rates were calculated using the 2013 European Standard Population aged 40-80 years.[16](https://webarchive.nationalarchives.gov.uk/ukgwa/20160105231503/http://www.ons.gov.uk/ons/guide-method/user-guidance/health-and-life-events/index.html) Rates are presented per 100 000 person years at risk (supplementary methods). - Briggs2022

>  Cancer Incidence Calculation 
Whole UKB cohort CRC incidence rates were calculated based on linked registry cases, without removal of prevalent cases, to reflect registration as would occur in national data. In addition ASIRs were calculated in the Integrated Modelling Cohort, in which prevalent cases were removed and cases identified through cancer and death registry, and linked hospital inpatient data; follow-up duration was as defined in the main methods. This analysis used R packages ‘survival’ and ‘epitools’. - [Briggs et al., 2022, p. 4]

The tables from Briggs2022 citations 15 and 16 are saved at K:\TEU\PRSforBrCa\Resources\referencetablearv2013c44_tcm77-410534.xls and K:\TEU\PRSforBrCa\Resources\agestandardisedmortalityratecalculationtemplateusingthe2013esp_tcm77-359944.xls respectively.

## UK Biobank overall

Prevalent and incident cases are included.

```{r}

# UKB overall (ie including prevalent cases)

# Our TEU_BrCa_status and TEU_BrCa_age are generated for prevalent cases
# So supplement with BrCaAge_Any (age at brca dx from Cancer Registry) to include the incident cases
# Note TEU_BrCa_age is the age at which follow-up ended, ie either breast cancer dx or censoring (death by other causes, lost to followup, admin censoring)

data <- wholepop %>% transmute(
  TEU_BrCa_status = as.numeric(!is.na(BrCaAge_Any) | TEU_BrCa_status), # Include prevalent cases
  TEU_BrCa_age = coalesce(BrCaAge_Any, TEU_BrCa_age), # Take CaR recorded age preferentially if it exists
  TEU_BaC_AgeAtRec = TEU_BaC_AgeAtRec
) %>%
  filter(TEU_BrCa_age >= 40) # Only interested in cancers from age 40

groups <- c()
brca_n <- c()
atrisk_n <- c()

for(age in c(55, 60, 65, 70, 75, 80)){

  # Define the 5-yr interval of interest
  start <- age
  end <- age + 5

  # Prep labels
  group <- paste0("[", start, ",", end, ")")
  groups <- c(groups, group)

  # Number of cases in this age group (had breast cancer and age at dx within this interval)
  brca <- data %>%
    filter(TEU_BrCa_status == 1 & TEU_BrCa_age >= start & TEU_BrCa_age < end)
  brca_n <- c(brca_n, nrow(brca))

  # Number of person years at risk in this age group
  atrisk <- data %>% 
    filter(TEU_BrCa_age >= start) %>% # If they weren't censored before the start of this interval
    mutate(start_int = start,
           end_int = pmin(TEU_BrCa_age, end),
           time_at_risk = end_int - start_int
           )

  atrisk_n <- c(atrisk_n, sum(atrisk$time_at_risk))
}

rates_wholepop <- data.frame(Age5 = groups,
                    brca = brca_n,
                    n = atrisk_n,
                    esp_2013 = c(6500, 6000, 5500, 5000, 4000, 2500)) %>%
  mutate(Age_specific_rates = brca/(n/100000),
         lci = 0.5 * qchisq(p=0.025, df=(2*brca))/(n/100000),
         uci = 0.5 * qchisq(p=0.975, df=(2*(brca+1)))/(n/100000),
         std_wt = esp_2013/sum(esp_2013),
         Age_standardised_rates = std_wt * Age_specific_rates)

sum(rates_wholepop$Age_standardised_rates)

kable(rates_wholepop, digits=2)

ASR <- sum(rates_wholepop$Age_standardised_rates)
LCI <- round(ASR - 1.96*(ASR/sqrt(sum(rates_wholepop$brca))), 2)
UCI <- round(ASR + 1.96*(ASR/sqrt(sum(rates_wholepop$brca))), 2)

```

The age-standardised breast cancer rate per 100,000 population (ages 55-85) is: `r round(ASR,2)` (95% CI: `r LCI`, `r UCI`).

This compares to the ONS 2013 directly age-standardised rates per 100,000 population of 299.5 for women.

## Analysis population

Population is post-menopausal White British women with no history of breast cancer, carcinoma in situ or risk reducing mastectomy.

Only incident cases are considered.

```{r}

# Analysis population

data <- analysis_pop %>% select(
  TEU_BrCa_status,
  TEU_BrCa_age,
  TEU_BaC_AgeAtRec
)

groups <- c()
brca_n <- c()
atrisk_n <- c()

for(age in c(55, 60, 65, 70, 75, 80)){

  start <- age
  end <- age + 5

  group <- paste0("[", start, ",", end, ")")
  groups <- c(groups, group)

  # Number of cases in this age group (had breast cancer and age at dx within this interval)
  brca <- data %>% 
    filter(TEU_BrCa_status == 1 & TEU_BrCa_age >= start & TEU_BrCa_age < end)
  brca_n <- c(brca_n, nrow(brca))

  # Number of person years at risk in this age group
  atrisk <- data %>% 
    filter(TEU_BaC_AgeAtRec < end & TEU_BrCa_age >= start) %>% # Risk only starts at baseline in analysis pop
    mutate(start = pmax(TEU_BaC_AgeAtRec, start),
           end = pmin(TEU_BrCa_age, end),
           time_at_risk = end - start
           )

  atrisk_n <- c(atrisk_n, sum(atrisk$time_at_risk))
}

rates_analysispop <- data.frame(Age5 = groups,
                    brca = brca_n,
                    n = atrisk_n,
                    esp_2013 = c(6500, 6000, 5500, 5000, 4000, 2500)) %>%
  mutate(Age_specific_rates = brca/(n/100000),
         lci = 0.5 * qchisq(p=0.025, df=(2*brca))/(n/100000),
         uci = 0.5 * qchisq(p=0.975, df=(2*(brca+1)))/(n/100000),
         std_wt = esp_2013/sum(esp_2013),
         Age_standardised_rates = std_wt * Age_specific_rates)

sum(rates_analysispop$Age_standardised_rates)

kable(rates_analysispop, digits=2)

ASR <- sum(rates_analysispop$Age_standardised_rates)
LCI <- round(ASR - 1.96*(ASR/sqrt(sum(rates_analysispop$brca))), 2)
UCI <- round(ASR + 1.96*(ASR/sqrt(sum(rates_analysispop$brca))), 2)

```

The age-standardised breast cancer rate per 100,000 population (ages 55-85) is: `r round(ASR,2)` (95% CI: `r LCI`, `r UCI`).

This compares to the ONS directly age-standardised rates per 100,000 population of 299.5 for women.

# Plot age-specific breast cancer rates, compare to ONS

```{r}

# Plot results from
# Analysis population: above chunk
# Whole UKB population: branch wholepop
# ONS 2013 data: K:\TEU\PRSforBrCa\Resources\referencetablearv2013c44_tcm77-410534.xls

rateplot <- cbind(rates_analysispop %>% transmute(Age5 = Age5, 
                                                  rate_analysis = Age_specific_rates,
                                                  lci_analysis = lci,
                                                  uci_analysis = uci),
                  rates_wholepop %>% transmute(rate_wholeUKB = Age_specific_rates,
                                               lci_wholeUKB = lci,
                                               uci_wholeUKB = uci),
               rate_ONS = c(276.0, 354.8, 415.9, 354.9, 400.6, 432.1)
               ) %>%
  mutate(lci_ONS = rate_ONS,
         uci_ONS = rate_ONS) %>%
  pivot_longer(cols=c(rate_analysis, rate_wholeUKB, rate_ONS,
                      lci_analysis, lci_wholeUKB, lci_ONS,
                      uci_analysis, uci_wholeUKB, uci_ONS),
               names_to = c(".value", "name"),
               names_sep = "_") %>%
  mutate(Cohort = factor(name, levels=c("ONS", "wholeUKB", "analysis"),
                       c("ONS 2013", "Whole UK Biobank", "Analysis population")))

ggplot(data=rateplot, aes(x=Age5, y=rate, group=Cohort, color=Cohort)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=lci, ymax=uci), width=.3, position=position_dodge(0.1)) +
  ggtitle("Age specific breast cancer rates in women in the UK Biobank,\noverall and within analysis cohort, compared to\nOffice for National Statistics 2013 Cancer Registry data") +
  ylab("Incidence rate per 100,000 person years at risk") +
  xlab("Age at cancer diagnosis, years")

# ggsave(file.path(config$outputs$figures, "incidence_rates.jpg"), 
#        scale=0.7, width=25, height=15, units="cm")

```

# Yearly incidence rates in UKB as sanity check

```{r}

data_wholepop <- wholepop %>% transmute(
  TEU_BrCa_status = as.numeric(!is.na(BrCaAge_Any) | TEU_BrCa_status),
  TEU_BrCa_age = coalesce(BrCaAge_Any, TEU_BrCa_age), # Take CaR recorded age if exists
  TEU_BaC_AgeAtRec = TEU_BaC_AgeAtRec
) %>%
  filter(TEU_BrCa_age >= 40) # Making this explicit

data_analysis <- analysis_pop %>% select(
  TEU_BrCa_status,
  TEU_BrCa_age,
  TEU_BaC_AgeAtRec
)

# Yearly incidence
cases_wholepop <- c()
cases_analysispop <- c()
people_wholepop <- c()
people_analysispop <- c()

for(age in seq(40, 84, by=1)){

 cases_wholepop <- c(cases_wholepop, nrow(data_wholepop %>% filter(TEU_BrCa_status == 1 & round(TEU_BrCa_age) == age)))
 cases_analysispop <- c(cases_analysispop, nrow(data_analysis %>% filter(TEU_BrCa_status == 1 & round(TEU_BrCa_age) == age)))

 people_wholepop <- c(people_wholepop, nrow(data_wholepop %>% filter(TEU_BrCa_age >= age)))
 people_analysispop <- c(people_analysispop, nrow(data_analysis %>% filter(TEU_BaC_AgeAtRec <= age & TEU_BrCa_age >= age)))
}

rates <- data.frame(age = seq(40, 84, by=1),
                    cases_wholepop = cases_wholepop,
                    cases_analysispop = cases_analysispop,
                    people_wholepop = people_wholepop,
                    people_analysispop = people_analysispop) %>%
  mutate(rates_wholepop = cases_wholepop/(people_wholepop/100000),
         rates_analysispop = cases_analysispop/(people_analysispop/100000),
         rates_ons = rep(c(125.3, 230.4, 281.0, 276.0, 354.8, 415.9, 354.9, 400.6, 432.1), each=5)) %>%
  pivot_longer(cols=c(rates_analysispop, rates_wholepop, rates_ons)) %>%
  mutate(Cohort = factor(name, levels=c("rates_ons", "rates_wholepop", "rates_analysispop"),
                       c("ONS 2013", "Whole UK Biobank", "Analysis population")))

ggplot(data=rates, aes(x=age, y=value, group=Cohort, color=Cohort)) +
  geom_line() +
  geom_point() +
  ggtitle("Age specific breast cancer rates in women in the UK Biobank,\noverall and within analysis cohort, compared to\nOffice for National Statistics 2013 Cancer Registry data") +
  ylab("Incidence rate per 100,000 person years at risk") +
  xlab("Age at cancer diagnosis, years")

```

### 2023-11-22 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Initial commit

---
title: "Manuscript_Outputs"
author: "Jennifer Collister"
date: "`r Sys.Date()`"
output:
    word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE, message = FALSE)

library(kableExtra)
# Issue with kableExtra collapse_rows for data dictionary
# Use development version
# renv::install("haozhu233/kableExtra@a6af5c0")
# See https://github.com/haozhu233/kableExtra/issues/595#issuecomment-808703159

library(knitr)
library(yaml)
library(here)
library(ggplot2)
library(tidyr)
library(readr)
library(dplyr)
library(glue)
library(tictoc)

# Pretty plots
library(etm)
library(plotROC) # For XML dependency, used install.packages("XML", type = "binary") to avoid compilation issues
library(gridExtra)
library(maps)
library(pROC)

# Pretty tables
library(arsenal)
library(pander)
library(flextable)

# Survival 
library(survival)
library(survminer)
library(rms) # For Hmisc dependency, used install.packages("Hmisc@4.7-2") to avoid installation issues
library(lmtest)

# NRI
library(nricens)

# Gail model
library(BCRA)

# Flowchart
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)
library(png)

# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(file.path(config$functions)))

#-------------------------------------------------------------------------------------------------------------------

# Output Formatting

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)

# Pander options
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')

# Set formatting options for tableby objects
my_controls <- tableby.control(
    test = F,
    total = T,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "Nmiss"), #"medianq1q3", 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      # medianq1q3 = "Median (IQR)",
      # range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

by_status <- tableby.control(
    test = T,
    total = T,
    numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "Nmiss"), #"medianq1q3", 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      # medianq1q3 = "Median (IQR)",
      # range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

#-------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r derive-data, include=FALSE}

source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data) {

  excl$initial <<- nrow(data)

  # Restrict to females (by self-report and genetic)
  data <- data[data$BaC_Sex == "Female",]
  data <- data[data$GeP_Sex == "Female" & !is.na(data$GeP_Sex),]
  excl$sex <<- nrow(data)

  # Restrict to post-menopausal women
  data <- data[!is.na(data$FSF_MenopauseStatus) & data$FSF_MenopauseStatus == "Yes",]
  excl$meno <<- nrow(data)

  # Restrict to White British (genetically)
  data <- data[!is.na(data$GeP_ethnic),]
  excl$whitebrit <<- nrow(data)

  # Exclude those outside the 40-70 age range
  data <- data[!is.na(data$TEU_BaC_AgeAtRec) & data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  excl$agerange <<- nrow(data)
  # Restrict population to those for whom we have PRS scores available
  data <- data[!is.na(data$PRS_bc_std),]
  excl$prs <<- nrow(data)
  # Exclude women with 20+ sisters as the TC model can't handle it
  data <- data[is.na(data$TC_NSisters) | data$TC_NSisters < 20,]
  # Exclude those who didn't answer HRT use
  data <- data[data$TC_HRT_raw != "Unanswered",]
  # Or who reported impossible HRT dates
  data <- data[is.na(data$FSF_HRTAgeStop) | data$FSF_HRTAgeStop <= ceiling(data$TEU_BaC_AgeAtRec),]
  excl$hrt <<- nrow(data)

  # Exclude prior breast cancer
  data <- data[is.na(data$BrCaDx_Prevalent),]
  data <- data[is.na(data$VI_BrCa),]
  excl$brca <<- nrow(data)
  # Prior mastectomy
  data <- data[data$TEU_HES_Mast_prev == "No",]
  excl$mast <<- nrow(data)
  # Prior carcinoma in situ
  data <- data[is.na(data$BrCaInSitu_Prevalent),]
  excl$cainsitu <<- nrow(data)

  return(data)
}

excl <- list(initial=0)
tic("Data derived: ")
data <- derive_variables(
  database = config$data$database,
  field_definitions = TEU_SPECS$PRSforBrCa,
  exclusions = exclusions,
  dictionary = "PRSforBrCa.html"
)
toc()

pretty_names <- pretty_switch(field_definitions=TEU_SPECS$PRSforBrCa)
pretty_list <- pretty_switch(field_definitions=TEU_SPECS$PRSforBrCa, return_type="list")

# Standardise variables so HR are per sd
data <- data %>% mutate(
  TC_10risk_sd = TC_10risk_cal/sd(TC_10risk_cal),
  gail_10risk_sd = gail_10risk_cal/sd(gail_10risk_cal),
  PRS_bc_sd = PRS_bc_std/sd(PRS_bc_std),
  TEU_BrCa_313_PRS_center = TEU_BrCa_313_PRS_center / sd(TEU_BrCa_313_PRS_center),
  TEU_BrCa_100k_PRS_center = TEU_BrCa_100k_PRS_center / sd(TEU_BrCa_100k_PRS_center)
  )

# Separate out test and training data
# train_IDs <- data$ID[data$train_set==TRUE]
# write.csv(train_IDs, file.path(config$data$derived, "training_ids.csv"), row.names = FALSE)
train_IDs <- read.csv(file.path(config$data$derived, "training_ids.csv"))$x

train_data <- data %>% filter(ID %in% train_IDs)

test_data <- data %>% filter(!ID %in% train_IDs)

```

```{r}
label1 <- paste0('UKB participants\\nn = ', excl$initial)
label2 <- paste0('Post-menopausal\nWhite British\nfemale participants\\nn = ', excl$agerange)
label3 <- paste0('Aged between 40 and 70\nwith available core data\\nn = ', excl$hrt)
label4 <- paste0('Study population: post-menopausal women\nwith no history of breast cancer\\nn = ', excl$cainsitu)

label5 <- paste0(excl$initial-excl$sex, ' males or genetically sex discordant\\l', excl$sex-excl$meno, ' pre-menopausal\\lor unknown menopausal status \\l', excl$meno-excl$whitebrit, ' not of genetically White British ancestry\\l')
label6 <- paste0(excl$whitebrit-excl$agerange,' outside the age range of [40,70) years \\l', excl$agerange-excl$prs,' Genomics plc PRS unavailable\\l', excl$prs-excl$hrt, ' invalid or missing data for core variables*\\l')
label7 <- paste0(excl$hrt-excl$brca, ' previous breast cancer\\l', excl$brca-excl$mast, ' previous mastectomy\\l', excl$mast-excl$cainsitu,' previous carcinoma in situ\\l')

gv <- read_file(here::here(file.path(config$outputs$flowcharts, "Exclusions.gv")))

export_svg(DiagrammeR::grViz(glue(gv))
           ) %>% charToRaw %>% rsvg %>%
  png::writePNG(here::here(file.path(config$outputs$figures, "Exclusions.png")))
```

## Cox model containing breast cancer risk score and PRS in training data (N=`r nrow(train_data)`)

```{r, fig.height=6}
# Incorporate PRS using a Cox model in training data, then predict 10-yr risks in the test data
plot_varnames <- function(x) {
  y_names <- dplyr::case_when(
    x == "TC_10risk_sd" ~ "Tyrer-Cuzick 10yr risk",
    x == "gail_10risk_sd" ~ "Gail 10yr risk",
    x == "PRS_bc_sd" ~ "PRS_BC",
    TRUE ~ x
  )
  return(y_names)
}

# Mimic get.risk.coxph from nricens but modify to use prediction in new data
get.risk.coxph_test <- function(mdl, t0, newdata) {
  bash <- basehaz(mdl)
  lambda0 <- approx(bash$time, bash$hazard, t0)$y
  linpred <- predict(mdl, newdata=newdata, type="lp")
  risk <- 1 - exp(-lambda0 * exp(linpred))
}

# Tyrer-Cuzick
model_TC <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        TC_10risk_sd + PRS_bc_sd + GeP_Array + 
                        GPLC_PC_1 + GPLC_PC_2 + GPLC_PC_3 + GPLC_PC_4,
                 data=train_data)

test_data$TC_prs_10risk <- get.risk.coxph_test(model_TC, t0=10, newdata=test_data)
# train_data$TC_prs_10risk2 <- predict(model_TC, type = "risk")

# plot Schoenfeld residuals
test.ph <- cox.zph(model_TC)
# test.ph
y <- test.ph$y
colnames(y) <- plot_varnames(colnames(y))
test.ph$y <- y
ggcoxzph(test.ph, point.alpha=0.04, var=c("Tyrer-Cuzick 10yr risk", "PRS_BC"),
         caption="Model also includes genetic array and first 4 principal components of genetic ancestry, plots omitted.")

# Gail
model_gail <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        gail_10risk_sd + PRS_bc_sd + GeP_Array + 
                        GPLC_PC_1 + GPLC_PC_2 + GPLC_PC_3 + GPLC_PC_4,
                 data=train_data)

test_data$gail_prs_10risk <- get.risk.coxph_test(model_gail, t0=10, newdata=test_data)

# plot Schoenfeld residuals
test.ph <- cox.zph(model_gail)
# test.ph
y <- test.ph$y
colnames(y) <- plot_varnames(colnames(y))
test.ph$y <- y
ggcoxzph(test.ph, point.alpha=0.04, var=c("Gail 10yr risk", "PRS_BC"),
         caption="Model also includes genetic array and first 4 principal components of genetic ancestry, plots omitted.")

# Test data

test_data <- test_data %>% 
  mutate(
    TC_10risk_cat = factor(ifelse(TC_10risk_cal <= 0.05, "<=5%", ">5%"),
                         levels=c("<=5%", ">5%")),
    gail_10risk_cat = factor(ifelse(gail_10risk_cal <= 0.05, "<=5%", ">5%"),
                           levels=c("<=5%", ">5%")),
    TC_prs_10risk_cat = factor(ifelse(TC_prs_10risk <= 0.05, "<=5%", ">5%"),
                             levels=c("<=5%", ">5%")),
  gail_prs_10risk_cat = factor(ifelse(gail_prs_10risk <= 0.05, "<=5%", ">5%"),
                               levels=c("<=5%", ">5%"))
  )

# Table
get_HR_perSD <- function(model, var) {

  coeff <- summary(model)$coefficients
  HR <- coeff[rownames(coeff)==var,]

  return(paste0(pretty_dp(exp(HR[1]), dp=2), " ",
                pretty_confint(exp(HR[1] - (1.96*HR[3])), 
                               exp(HR[1] + (1.96*HR[3])), dp=2))
  )
}

dt <- data.frame(var=c("Tyrer-Cuzick", "Gail"),
                 model=c(get_HR_perSD(model_TC, "TC_10risk_sd"), get_HR_perSD(model_gail, "gail_10risk_sd")),
                 prs=c(get_HR_perSD(model_TC, "PRS_bc_sd"), get_HR_perSD(model_gail, "PRS_bc_sd"))
)

ft <- flextable(dt) %>%
  set_header_labels(var="Model", model="10-year risk score HR per SD", prs="PRS HR per SD") %>%
  set_table_properties(layout="autofit") %>%
  footnote(i=1, j=1, 
           value=as_paragraph("Hazard ratio per standard deviation from a Cox model containing 10-year risk score from breast cancer risk prediction model, Genomics plc PRS, genetic array and first 4 principal components of genetic ancestry, in the training data"), 
           ref_symbols="a", part="header")

ft

```

```{r}
# How many women are above 5% 10yr risk according to the original model?

TC_5pct <- table(test_data$TC_10risk_cal > 0.05, useNA='ifany')
TC_5prop <- TC_5pct[2]/nrow(test_data)
TC_prs_threshold <- quantile(test_data$TC_prs_10risk, probs=c(1-TC_5prop))

gail_5pct <- table(test_data$gail_10risk_cal > 0.05, useNA='ifany')
gail_5prop <- gail_5pct[2]/nrow(test_data)
gail_prs_threshold <- quantile(test_data$gail_prs_10risk, probs=c(1-gail_5prop))

# How many women are above 5% 10yr risk after adding the PRS?

TC_combined_5pct <- table(test_data$TC_prs_10risk > 0.05)

gail_combined_5pct <- table(test_data$gail_prs_10risk > 0.05)

test_data <- test_data %>% mutate(
  TC_prs_10risk_adj = TC_prs_10risk - (quantile(TC_prs_10risk, probs=c(1-TC_5prop)) - 0.05),
  TC_prs_10risk_adj_cat = factor(ifelse(TC_prs_10risk <= TC_prs_threshold, 
                                          paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                          paste0(">", round(100*TC_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                        paste0(">", round(100*TC_prs_threshold,2), "%"))),
  gail_prs_10risk_adj = gail_prs_10risk - (quantile(gail_prs_10risk, probs=c(1-gail_5prop)) - 0.05),
  gail_prs_10risk_adj_cat = factor(ifelse(gail_prs_10risk <= gail_prs_threshold, 
                                          paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                          paste0(">", round(100*gail_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                        paste0(">", round(100*gail_prs_threshold,2), "%")))
  )

```

```{r, results=FALSE}

# Net reclassification index, for Table 1

# Tyrer-Cuzick
nri_tc <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_prs_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

# Gail
nri_gail <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$gail_10risk_cal, p.new=test_data$gail_prs_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

nri_ci <- function(nri_output, type="overall", dp=3){
  row <- dplyr::case_when(
    type=="overall" ~ 1,
    type=="case" ~ 2,
    type=="control" ~ 3
  )
  est <- nri_output$nri$Estimate[row]
  lci <- nri_output$nri$Lower[row]
  uci <- nri_output$nri$Upper[row]

  return(paste0(pretty_dp(est, dp=dp), " ", pretty_confint(lci, uci, dp=dp)))
}

```

## Table `r table`: Model performance with/without PRS in test data (N=`r nrow(test_data)`)
`r table <- table + 1`

```{r}

dt <- data.frame(model=c("Tyrer-Cuzick", "Model only", "Model with PRS", 
                         "Gail", "Model only", "Model with PRS"),
                 AUC=c(NA,
                       AUC_CI(df=test_data, score="TC_10risk_cal", status = "TEU_BrCa_status", dp=2),
                       AUC_CI(df=test_data, score="TC_prs_10risk", status = "TEU_BrCa_status", dp=2),
                       NA,
                       AUC_CI(df=test_data, score="gail_10risk_cal", status = "TEU_BrCa_status", dp=2),
                       AUC_CI(df=test_data, score="gail_prs_10risk", status = "TEU_BrCa_status", dp=2)),
                 NRI_overall=c(NA, nri_ci(nri_tc, type="overall", dp=3), nri_ci(nri_tc, type="overall", dp=3), 
                               NA, nri_ci(nri_gail, type="overall", dp=3), nri_ci(nri_gail, type="overall", dp=3)),
                 NRI_case=c(NA, nri_ci(nri_tc, type="case", dp=3), nri_ci(nri_tc, type="case", dp=3), 
                               NA, nri_ci(nri_gail, type="case", dp=3), nri_ci(nri_gail, type="case", dp=3)),
                 NRI_control=c(NA, nri_ci(nri_tc, type="control", dp=3), nri_ci(nri_tc, type="control", dp=3), 
                               NA, nri_ci(nri_gail, type="control", dp=3), nri_ci(nri_gail, type="control", dp=3)))

ft <- flextable(dt) %>%
  merge_v(j = ~ model + NRI_overall + NRI_case + NRI_control) %>%
  # separate_header(sep="_") %>%
  set_header_labels(model="Model", AUC="AUC (95% CI)",
                    NRI_overall="Overall", 
                    NRI_case=paste0("Case (N=", nrow(test_data[test_data$TEU_BrCa_time <= 10 &
                                                                 test_data$TEU_BrCa_status==1,]), ")"), 
                    NRI_control=paste0("Control (N=", nrow(test_data[test_data$TEU_BrCa_time > 10,]), ")")) %>%
  add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
  align(align = "center", part = "all", j=c('AUC', 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
  bg(i= ~ is.na(AUC), bg="lightgrey", part="body") %>%
  merge_at(i=1, j=1:5) %>%
  merge_at(i=4, j=1:5) %>%
  fix_border_issues(part="all") %>%
  set_table_properties(layout="autofit") %>%
  footnote(i=1, j=3, 
           value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk"), 
           ref_symbols="a", part="header") %>%
  footnote(i=2, j=3, 
           value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up."), 
           ref_symbols="b", part="header")

ft

# Flextable: Border issues with merged cells
# https://davidgohel.github.io/flextable/reference/fix_border_issues.html

```

## Figure `r figure`: Calibration plots
`r figure <- figure + 1`

```{r, fig.height=7}

# Calibration - code adapted from Ruthie's Project and Xiaonan's diabetes work with Leicester score

cal_plot <- function(data, var="TC_10risk", title, xlim=c(0.01,0.07), ylim=c(0.01,0.07)){

  # First group into deciles of predicted risk
  deciles <- quantile(data[[var]], probs = seq(0, 1, 0.1))
  labels <- c("1st decile", "2nd", "3rd", "4th", "5th", "6th", "7th", "8th", "9th", "10th decile")
  data <- data %>% mutate(
    TEU_BrCa_status = ifelse(TEU_BrCa_time < 10, TEU_BrCa_status, 0),
    TEU_BrCa_time = ifelse(TEU_BrCa_time < 10, TEU_BrCa_time, 10),
    risk_dec = cut(get(var), breaks=deciles, labels=labels, right=TRUE, include.lowest=TRUE)
    )

  # Mean predicted risk per decile
  predicted <- data %>%
    group_by(risk_dec) %>% 
    summarise(predicted = mean(get(var)))

  # Fit survival curves from Kaplan-Meier and extract observed risks by predicted risk decile
  fit <- survfit(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ risk_dec, data=data)
  fit_time <- summary(fit, times = 10)

  observed <- data.frame(risk_dec = fit_time$strata,
                         observed = 1 - fit_time$surv,
                         observed_lci = 1 - fit_time$lower,
                         observed_uci = 1 - fit_time$upper) %>%
    mutate(risk_dec = as.factor(str_remove(risk_dec, "risk_dec=")))

  # Join predicted and observed
  calibration <- predicted %>% inner_join(observed, by="risk_dec")

  # Calculate calibration slope
  cal_model <- lm(observed ~ predicted, data=calibration)
  cal_slope <- cal_model$coefficients[2]
  cal_intercept <- cal_model$coefficients[1]

  # Plot in calibration plot
  plot <- ggplot(calibration, aes(predicted, observed)) +
    geom_point(shape=22, size=3, color="blue", fill="blue") +
    geom_abline(aes(slope=1, intercept=0, linetype="dashed", colour="Ideal (y=x)"), show.legend=FALSE) +
    geom_smooth(method=lm, aes(colour="Best fit")) +
    geom_pointrange(aes(ymin=observed_lci, ymax=observed_uci)) +
    scale_x_continuous(limits=xlim) +
    scale_y_continuous(limits=ylim) +
    labs(x="Predicted risk",
         y="Observed risk",
         title=title) +
    annotate("text", x=0.05, y = 0.02, size=10/.pt,
             label=paste0("Slope = ", pretty_dp(cal_slope, 3), 
                          "\nIntercept = ", pretty_dp(cal_intercept, 3))) +
    scale_color_manual(name = "Legend",             # legend name
                     values = c("Ideal (y=x)" = "black",  # map regression line colors
                                "Best fit" = "blue")) +
    theme(legend.justification = c(0, 1), legend.position = c(0, 1)) +
    theme(plot.title = element_text(size=10))

  ggsave(file.path(config$outputs$figures, paste0(var, "_calibration_plot10.jpg")), 
       scale=0.6, width=25, height=15, units="cm")

  return(plot)

}

TC_cal <- cal_plot(data=data, var="TC_10risk",
                   title="Tyrer-Cuzick model")
gail_cal <- cal_plot(data=data, var="gail_10risk",
                     title="Gail model") +
  theme(legend.position="none")
TC_cal2 <- cal_plot(data=data, var="TC_10risk_cal",
                   title="Calibrated Tyrer-Cuzick model")
gail_cal2 <- cal_plot(data=data, var="gail_10risk_cal",
                     title="Calibrated Gail model") +
  theme(legend.position="none")

# 2. Save the legend
#+++++++++++++++++++++++
legend <- get_legend(TC_cal)
# 3. Remove the legend from the box plot
#+++++++++++++++++++++++
TC_cal <- TC_cal + theme(legend.position="none")

gridExtra::grid.arrange(TC_cal, gail_cal, legend, TC_cal2, gail_cal2,
                        ncol=3, widths=c(3.5, 3.5, 0.8),
                        top=paste0("Before and after calibration plots of 10-year observed vs predicted risk by predicted risk decile of each model in the whole analysis population (N=", nrow(data), ")"))

```

## Figure `r figure`: ROC Plots
`r figure <- figure + 1`

```{r, fig.width=12, warning=FALSE}

# ROC plot
# https://rviews.rstudio.com/2019/03/01/some-r-packages-for-roc-curves/
# https://cran.r-project.org/web/packages/plotROC/plotROC.pdf
# https://cran.r-project.org/web/packages/plotROC/vignettes/examples.html

# New approach uses pROC, Xiaonan's code
# https://gitlab.ndph.ox.ac.uk/jenniferco/prsondiabetes/-/blob/Riskscore/Scripts/Stats_Analysis/RMarkdown/RiskScore_Primary.Rmd#auc

roc_plot <- function(data, modelname, prsname, plotname,
                     varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk")){

  AUC_CIs <- sapply(varlist, function(x) AUC_CI(df=data, score=x, status = "TEU_BrCa_status", dp=2))

  AUC_labels <- c(
    paste0(modelname, " 10-yr risk: ", AUC_CIs[1]),
    paste0(prsname, " PRS: ", AUC_CIs[2]),
    paste0(modelname, " & PRS 10-yr risk: ", AUC_CIs[3])
  )

  long <- data %>%
    select(ID, TEU_BrCa_status, all_of(varlist)) %>%
    pivot_longer(!c("ID","TEU_BrCa_status"), names_to = "Score", values_to = "Value") %>%
    mutate(Score=factor(Score,levels = varlist))

  plot <- ggplot(long, aes(d=TEU_BrCa_status, m=Value, color=Score)) +
    geom_roc(n.cuts = 0, linealpha = 1, size=1) +
    style_roc() +
    geom_abline(aes(slope=1, intercept=0), linetype="dashed") +
    scale_color_manual(name="AUC (95% CI)", labels =  AUC_labels, values = c("#F8766D","#00BA38", "#6193FF")) +
    labs(title = paste0(plotname), 
         caption = "Dashed line is y=x") +
    theme(legend.position = c(0.7, 0.15))

  ggsave(file.path(config$outputs$figures, paste0("rocplot_", modelname, ".jpg")), 
         scale=0.7, width=25, height=15, units="cm")

  plot
}

# Tyrer-Cuzick
TC_roc <- roc_plot(test_data, modelname="TC", prsname="Genomics plc", plotname="Tyrer-Cuzick",
         varlist=c("TC_10risk_cal", "PRS_bc_std", "TC_prs_10risk"))

# Gail
gail_roc <- roc_plot(test_data, modelname="Gail", prsname="Genomics plc", plotname="Gail",
         varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk"))

gridExtra::grid.arrange(TC_roc, gail_roc,
                        ncol=2,
                        top=paste0("ROC curves for 10-yr risk predictions in test data (N=", nrow(test_data), ")"))

```

## Tables 3 Reclassification

### Tyrer-Cuzick

When using the Tyrer-Cuzick model, `r pretty_dp(100*TC_5prop, 2)`% of women are above 5% 10yr risk.

This rises to `r pretty_dp(100*TC_combined_5pct[2]/nrow(test_data),2)`% of women above this 5% threshold when using the combination of both Tyrer-Cuzick and PRS in a Cox model.

Instead of using a fixed threshold, let's use the same proportion of women.

NRI using top `r pretty_dp(100*TC_prs_threshold,2)`% of 10yr risk as risk threshold

```{r}

nri_tab <- function(p, population, model_old, model_new) {
  p <- propped(p)
  p <- as.data.frame(cbind(c(model_old, model_old, ""), rownames(p), p))

  flextable(p) %>%
    merge_v(j = ~ V1) %>%
    set_header_labels(V1="", V2="") %>%
    add_header_row(values=c(population, model_new, ""), colwidths=c(2, 2,1)) %>%
    align(align = "center", part = "all") %>%
    set_table_properties(layout="autofit") %>%
    border_remove() %>%
    hline_top() %>%
    hline(i=c(2)) %>%
    vline(j=c(2, 4)) %>%
    fix_border_issues(part="all") %>%
    footnote(i=1, j=1, value=as_paragraph("Cases defined as individuals diagnosed with breast cancer within 10 years. Controls defined as individuals who were still at risk of breast cancer by 10 years of follow-up. Individuals censored before 10 years are not displayed."), 
           ref_symbols="a", part="header")
}

################
# Tyrer-Cuzick #
################

# Cases
print("Cases")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1], 
              test_data$TC_prs_10risk_adj_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]), 
        population="Cases", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time > 10], 
              test_data$TC_prs_10risk_adj_cat[test_data$TEU_BrCa_time > 10]), 
        population="Controls", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

```

### Gail

When using the Gail model, `r round(100*gail_5prop, 2)`% of women are above 5% 10yr risk.

This rises to `r round(100*gail_combined_5pct[2]/nrow(test_data),2)`% of women above this 5% threshold when using the combination of both Gail and PRS in a Cox model.

Instead of using a fixed threshold, let's use the same proportion of women.

NRI using top `r round(100*gail_prs_threshold,2)`% of 10yr risk as risk threshold

```{r}

########
# Gail #
########

# Cases
print("Cases")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1], 
              test_data$gail_prs_10risk_adj_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]), 
        population="Cases", model_old="Gail", model_new="Gail + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time > 10], 
              test_data$gail_prs_10risk_adj_cat[test_data$TEU_BrCa_time > 10]), 
        population="Controls", model_old="Gail", model_new="Gail + PRS")

```

# Decision curve analysis

```{r, fig.width=10}

library(dcurves)

name_conv <- list("Treat All" = "Screen all",
                  "Treat None" = "Screen none",
                  "TC_10risk_cal" = "Tyrer-Cuzick",
                  "TC_prs_10risk" = "Tyrer-Cuzick with PRS_BC",
                  "gail_10risk_cal" = "Gail",
                  "gail_prs_10risk" = "Gail with PRS_BC")
name_color <- c("Screen all" = "grey",
                "Screen none" = "black",
                "Tyrer-Cuzick" = "#E69F00",
                "Tyrer-Cuzick with PRS_BC" = "#D55E00",
                "Gail" = "#56B4E9",
                "Gail with PRS_BC" = "#0072B2")

allnone_drop <- function(.data) {
  new_dca <- .data$dca %>% 
    filter(! label %in% c("Treat None")) %>% 
    mutate(label = factor(label))
  .data$dca <- new_dca
  return(.data)
}

nameswap <- function(.data) {
  new_dca <- .data$dca %>% mutate(
    label = factor(label, labels=unlist(name_conv[levels(label)], use.names=FALSE))
    )
  .data$dca <- new_dca
  return(.data)
}

# Net benefit
dca_tc <- dca(TEU_BrCa_status ~ TC_10risk_cal + TC_prs_10risk, 
            test_data, 
            thresholds = seq(0, 0.12, by = 0.001)) %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Tyrer-Cuzick") + 
  theme(legend.position = c(0.6, 0.8)) + 
  scale_color_manual(values = name_color)

dca_tc
ggsave(file.path(config$outputs$figures, "TC_DCA.png"), 
       scale=0.7, width=25, height=15, units="cm")

dca_gail <- dca(TEU_BrCa_status ~ gail_10risk_cal + gail_prs_10risk,
            test_data,
            thresholds = seq(0, 0.12, by = 0.001)) %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Gail") +
  theme(legend.position = c(0.6, 0.8)) + 
  scale_color_manual(values = name_color)

dca_gail
ggsave(file.path(config$outputs$figures, "gail_DCA.png"), 
       scale=0.7, width=25, height=15, units="cm")

gridExtra::grid.arrange(dca_tc, dca_gail,
                        ncol=2,
                        top=paste0("Decision curves for 10-year risk from models with and without PRS_BC in test data (N=",
                                   nrow(test_data), ")"))

# Interventions avoided
ia_tc <- dca(TEU_BrCa_status ~ TC_10risk_cal + TC_prs_10risk, 
             data = test_data, 
             thresholds = seq(0, 0.12, by = 0.001),
             as_probability = c("TC_10risk_cal", "TC_prs_10risk")) %>%
  net_intervention_avoided(nper=100) %>%
  allnone_drop() %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Tyrer-Cuzick") + 
  theme(legend.position = c(0.3, 0.8)) + 
  scale_color_manual(values = name_color)

ia_tc
ggsave(file.path(config$outputs$figures, "TC_NIA.png"), 
       scale=0.7, width=25, height=15, units="cm")

ia_gail <- dca(TEU_BrCa_status ~ gail_10risk_cal + gail_prs_10risk,
               test_data,
               thresholds = seq(0, 0.12, by = 0.001)#,
               # as_probability = c("TC_10risk_cal", "TC_prs_10risk")
            ) %>%
  net_intervention_avoided(nper=100) %>%
  allnone_drop() %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Gail") +
  theme(legend.position = c(0.3, 0.8)) + 
  scale_color_manual(values = name_color)

ia_gail
ggsave(file.path(config$outputs$figures, "gail_NIA.png"), 
       scale=0.7, width=25, height=15, units="cm")

gridExtra::grid.arrange(ia_tc, ia_gail,
                        ncol=2,
                        top=paste0("Net interventions avoided per 100 women for 10-year risk from models with and without PRS_BC in test data (N=",
                                   nrow(test_data), ")"))

```

# Supplementary Materials

## Supplementary Figure: Flowchart illustrating selection of analytic dataset (n = `r nrow(data)`).

!["*'Invalid or missing data for core variables' comprises: 1. Missing data for HRT use (Never, Previous or Current) as this variable was required for the Tyrer-Cuzick model. Previous HRT users with unknown age of stopping HRT were categorised into 'Previous (<5)' and 'Previous (>5)' by imputation. 2. Twenty or more sisters, as this crashes the Tyrer-Cuzick model. 3. An age of stopping HRT that was greater than age at baseline."](Figures/Manuscript/Exclusions.png){height=700px}

## Supplementary Table: Baseline characteristics

Including variables from both models, by quintile (1st, 3rd, 5th, overall) of Genomics plcs PRS

```{r, results='asis'}

# Descrp stats
part1 <- tableby(PRS_bc_std_q ~ 
                   # Standard
                   TEU_BaC_AgeCat + Height + Weight +
                   # Reproductive
                   FSF_AgeMenarche + TC_Parity + FSF_AgeMenopause +
                   TC_HRT_imputed +
                   # Diagnoses
                   TC_OvarianCaDx + Biopsy_Prevalent +
                   # Family history
                   TC_MotherBrCa + 
                   TC_NSisters + TC_SiblingBrCa + TC_FatherBrCa,
                 data=data %>% mutate(Biopsy_Prevalent=factor(Biopsy_Prevalent)), 
                 control=my_controls)

# Descriptive stats within subgroups

# Reproductive
rep_agefb <- tableby(PRS_bc_std_q ~ AgeFirstBirth,
                 data=data %>% filter(!is.na(TC_Parity) & TC_Parity == "Parous"),
                 control=my_controls)

rep_hrtdur <- tableby(PRS_bc_std_q ~ FSF_HRTDuration,
                 data=data %>% filter(TC_HRT_imputed %in% c("Current",
                                                            "Previous user (less than 5 years ago)")),
                 control=my_controls)

rep_hrtsince <- tableby(PRS_bc_std_q ~ FSF_HRT_timesince,
                 data=data %>% filter(TC_HRT_imputed == "Previous user (less than 5 years ago)"),
                 control=my_controls)

# Medical history
dx_ov <- tableby(PRS_bc_std_q ~ TC_OvarianCaAge,
                 data=data %>% filter(TC_OvarianCaDx == "Yes"),
                 control=my_controls)

dx_hyp <- tableby(PRS_bc_std_q ~ Gail_AtypicalHyperplasia,
                  data=data %>% 
                    filter(Biopsy_Prevalent > 0) %>% 
                    mutate(Gail_AtypicalHyperplasia = factor(Gail_AtypicalHyperplasia)),
                  control=my_controls)

# Family history 
fh_ma <- tableby(PRS_bc_std_q ~ TC_MotherAge,
                 data=data %>% filter(TC_MotherBrCa == "No"),
                 control=my_controls)

summary(part1, labelTranslations = pretty_list,
        title=paste0('Table 1. Descriptive statistics by quintiles of breast cancer PRS'))

# Reproductive
summary(rep_agefb, labelTranslations = pretty_list,
        title="Table 1a. Age at first birth among parous women")
summary(rep_hrtdur, labelTranslations = pretty_list,
        title="Table 1b. Duration of HRT use, among women who are currently taking HRT or who previously used HRT less than 5 years ago")
summary(rep_hrtsince, labelTranslations = pretty_list,
        title="Table 1c. Time since stopping HRT among women who had previously used HRT less than 5 years ago")
# Medical history
summary(dx_ov, labelTranslations = pretty_list,
        title="Table 1d. Age at ovarian cancer diagnosis among women diagnosed with ovarian cancer")
summary(dx_hyp, labelTranslations = pretty_list,
        title="Table 1e. Diagnosis of atypical hyperplasia, among women with breast biopsy")
# Family history
summary(fh_ma, labelTranslations = pretty_list,
        title="Table 1f. Maternal age at baseline or death, among women whose mothers had not had breast cancer. Note: Tyrer-Cuzick model asked for maternal age at dx of breast cancer or, if no dx, then current age or age at death. In UKB data, maternal age at breast cancer diagnosis was not available, only age at baseline assessment or age at death.")

```

## Cox regression coefficients of individual variables in TC and Gail models

### TC

```{r, results='asis'}

# Tyrer-Cuzick

tc_varlist <- c("FSF_AgeMenopause", "FSF_AgeMenarche", "Height", "BMI", 
                "AgeFirstBirth", "TC_HRT_imputed", "AtypicalHyperplasia")
                # "TEU_BaC_AgeAtRec", "TC_Parity", 
                # "AtypicalHyperplasia", "TC_OvarianCaDx", "TC_OvarianCaAge", "FSF_HRTDuration", 
                # "FSF_HRT_timesince", "TC_MotherBrCa", "TC_NSisters", "TC_SiblingBrCa", "TC_FatherBrCa")

tc_data <- train_data %>% 
  mutate(BMI = Weight/(Height/100)^2,
         FSF_AgeMenopause = FSF_AgeMenopause/5,
         FSF_AgeMenarche = relevel(
           cut(FSF_AgeMenarche, breaks=c(0, 11, 12, 13, 14, 15, 16, 17, 100),
               labels=c("<11", "11", "12", "13", "14", "15", "16", "17 or older"), right=FALSE),
           ref="13"),
         Height = cut(Height, breaks=c(0, 160, 170, 200),
                      labels=c("<1.6", "1.6-1.7", "1.7 or taller"), right=FALSE),
         BMI = cut(BMI, breaks=c(0, 21, 23, 25, 27, 100),
                   labels=c("<21", "21 to <23", "23 to <25", "25 to <27", "27 or more"), right=FALSE),
         AgeFirstBirth = factor(dplyr::case_when(
           TC_Parity == "Nulliparous" ~ "Nulliparous",
           is.na(AgeFirstBirth) ~ NA,
           AgeFirstBirth < 20 ~ "<17-19",
           AgeFirstBirth < 25 ~ "20-24",
           AgeFirstBirth < 30 ~ "25-29",
           AgeFirstBirth < 35 ~ "30-34",
           AgeFirstBirth >= 35 ~ "35+"
         ), levels=c("Nulliparous", "<17-19", "20-24", "25-29", "30-34", "35+")),
         TC_HRT_imputed = factor(dplyr::case_when(
           TC_HRT_imputed == "Current" ~ "Current",
           TRUE ~ "Not current"
         ), levels=c("Not current", "Current"))) %>%
  select(ID, BaC_Age, TEU_BrCa_time, TEU_BrCa_status, 
                                 tc_varlist) %>%
  drop_na

tc_formula <- paste0("Surv(TEU_BrCa_time, TEU_BrCa_status) ~ BaC_Age + ", paste0(tc_varlist, collapse=" + "))
tc_cox <- coxph(as.formula(tc_formula), data=tc_data)

kable(printcoxresults(tc_data, varlist=tc_varlist, modeloutput=tc_cox))

for (var in tc_varlist) {
  tc_formula <- paste0("Surv(TEU_BrCa_time, TEU_BrCa_status) ~ BaC_Age + ", var)
  tc_cox <- coxph(as.formula(tc_formula), data=tc_data)
  print(kable(printcoxresults(tc_data, varlist=var, modeloutput=tc_cox)))
}

```

### Gail

```{r, results='asis'}

#  Gail

gail_varlist <- c("FSF_AgeMenarche", "Biopsy_Prevalent", "AgeCat",
                  "AgeFirstBirth", "FamilyBrCa_HowMany")

gail_data <- train_data %>% 
  mutate(Biopsy_Prevalent = factor(Biopsy_Prevalent),
         FSF_AgeMenarche = relevel(cut(FSF_AgeMenarche, breaks=c(0, 11, 13, 100),
               labels=c("<12", "12-13", ">=14"), right=TRUE), ref=">=14"),
         AgeFirstBirth = factor(dplyr::case_when(
           TC_Parity == "Nulliparous" ~ "25-29 or Nulliparous",
           is.na(AgeFirstBirth) ~ NA,
           AgeFirstBirth < 20 ~ "<20",
           AgeFirstBirth < 25 ~ "20-24",
           AgeFirstBirth < 30 ~ "25-29 or Nulliparous",
           AgeFirstBirth < 35 ~ "30 or older"
         ), levels=c("<20", "20-24", "25-29 or Nulliparous", "30 or older")),
         FamilyBrCa_HowMany = factor(FamilyBrCa_HowMany),
         AgeCat = cut(BaC_Age, breaks=c(0, 50, 100), labels=c("<50yr", ">=50yr"))
         ) %>%
  select(ID, BaC_Age, TEU_BrCa_time, TEU_BrCa_status, 
                                 gail_varlist) %>%
  drop_na

gail_cox <- coxph(as.formula(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ FSF_AgeMenarche + 
                   Biopsy_Prevalent + AgeCat + AgeFirstBirth*FamilyBrCa_HowMany), data=gail_data)
print(gail_cox)

gail_cox_nointeract <- coxph(as.formula(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ FSF_AgeMenarche + 
                   Biopsy_Prevalent + AgeCat + AgeFirstBirth + FamilyBrCa_HowMany), data=gail_data)

lmtest::lrtest(gail_cox, gail_cox_nointeract)

for (var in gail_varlist) {
  gail_formula <- paste0("Surv(TEU_BrCa_time, TEU_BrCa_status) ~ BaC_Age + ", var)
  gail_cox <- coxph(as.formula(gail_formula), data=gail_data)
  print(kable(printcoxresults(gail_data, varlist=var, modeloutput=gail_cox)))
}

```

# Distribution of risk

```{r}

facet_data <- test_data %>% select(TC_10risk_cal, TC_prs_10risk, TEU_BrCa_status) %>% 
  pivot_longer(cols=c(TC_10risk_cal, TC_prs_10risk))

density_data <- facet_data %>%
  mutate(
    name = ifelse(TEU_BrCa_status == 0, paste0(name, "_control"), paste0(name, "_case")),
    Model = factor(name, 
                   levels=c("TC_10risk_cal_control", "TC_10risk_cal_case",
                            "TC_prs_10risk_control", "TC_prs_10risk_case"),
                   labels = c("Tyrer-Cuzick controls", "Tyrer-Cuzick cases",
                              "Tyrer-Cuzick with PRS_BC controls", "Tyrer-Cuzick with PRS_BC cases")))

ggplot(density_data, 
       aes(x=value, fill=Model)) + 
  geom_density(alpha=0.4, position="identity") +
  ggtitle("Density plot of 10yr predicted risk by model in test data, split by cases and controls") +
  theme(legend.position = c(0.8, 0.8)) +
  xlim(0, 0.2) +
  geom_vline(xintercept=0.05)

ggsave(file.path(config$outputs$figures, "TC_density_10yrrisk.png"), 
       scale=0.7, width=25, height=15, units="cm")

```

```{r}

facet_data <- test_data %>% select(gail_10risk_cal, gail_prs_10risk, TEU_BrCa_status) %>% 
  pivot_longer(cols=c(gail_10risk_cal, gail_prs_10risk))

density_data <- facet_data %>%
  mutate(
    name = ifelse(TEU_BrCa_status == 0, paste0(name, "_control"), paste0(name, "_case")),
    Model = factor(name, 
                   levels=c("gail_10risk_cal_control", "gail_10risk_cal_case",
                            "gail_prs_10risk_control", "gail_prs_10risk_case"),
                   labels = c("Gail controls", "Gail cases",
                              "Gail with PRS_BC controls", "Gail with PRS_BC cases")))

ggplot(density_data, 
       aes(x=value, fill=Model)) + 
  geom_density(alpha=0.4, position="identity") +
  ggtitle("Density plot of 10yr predicted risk by model in test data, split by cases and controls") +
  theme(legend.position = c(0.8, 0.8)) +
  xlim(0, 0.2) +
  geom_vline(xintercept=0.05)

ggsave(file.path(config$outputs$figures, "Gail_density_10yrrisk.png"), 
       scale=0.7, width=25, height=15, units="cm")

```

# Fixed risk threshold

```{r}

nri <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_prs_10risk,
        t0=10, cut=0.05, msg=FALSE)

# kable(nri$nri)

################
# Tyrer-Cuzick #
################

# Cases
print("Cases")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1], 
              test_data$TC_prs_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]), 
        population="Cases", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time > 10], 
              test_data$TC_prs_10risk_cat[test_data$TEU_BrCa_time > 10]), 
        population="Controls", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

```

```{r}

nri <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$gail_10risk_cal, p.new=test_data$gail_prs_10risk,
        t0=10, cut=0.05, msg=FALSE)

# kable(nri$nri)

################
# Gail #
################

# Cases
print("Cases")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1],
              test_data$gail_prs_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]),
        population="Cases", model_old="Gail", model_new="Gail + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time > 10],
              test_data$gail_prs_10risk_cat[test_data$TEU_BrCa_time > 10]),
        population="Controls", model_old="Gail", model_new="Gail + PRS")

```

# Results using Mavaddat 313-SNP PRS

```{r, fig.height=9}

model_TC <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        TC_10risk_cal + TEU_BrCa_313_PRS_center + GeP_Array + 
                        GeP_PC_1 + GeP_PC_2 + GeP_PC_3 + GeP_PC_4,
                 data=train_data)

test_data$TC_prs313_10risk <- get.risk.coxph_test(model_TC, t0=10, newdata=test_data)

model_gail <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        gail_10risk_cal + TEU_BrCa_313_PRS_center + GeP_Array + 
                        GeP_PC_1 + GeP_PC_2 + GeP_PC_3 + GeP_PC_4,
                 data=train_data)

test_data$gail_prs313_10risk <- get.risk.coxph_test(model_gail, t0=10, newdata=test_data)

```

```{r}
# How many women are above 5% 10yr risk according to the original model?

TC_5pct <- table(test_data$TC_10risk_cal > 0.05, useNA='ifany')
TC_5prop <- TC_5pct[2]/nrow(test_data)
TC_prs_threshold <- quantile(test_data$TC_prs313_10risk, probs=c(1-TC_5prop))
TC_Mav2019_threshold <- quantile(test_data$TC_10risk_Mavaddat2019_cal, probs=c(1-TC_5prop))

gail_5pct <- table(test_data$gail_10risk_cal > 0.05, useNA='ifany')
gail_5prop <- gail_5pct[2]/nrow(test_data)
gail_prs_threshold <- quantile(test_data$gail_prs313_10risk, probs=c(1-gail_5prop))

# How many women are above 5% 10yr risk after adding the PRS?

TC_combined_5pct <- table(test_data$TC_prs313_10risk > 0.05)

gail_combined_5pct <- table(test_data$gail_prs313_10risk > 0.05)

test_data <- test_data %>% mutate(
  TC_prs313_10risk_adj = TC_prs313_10risk - (quantile(TC_prs313_10risk, probs=c(1-TC_5prop)) - 0.05),
  TC_prs313_10risk_adj_cat = factor(ifelse(TC_prs313_10risk <= TC_prs_threshold, 
                                          paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                          paste0(">", round(100*TC_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                        paste0(">", round(100*TC_prs_threshold,2), "%"))),
  TC_Mav2019_10risk_adj = TC_10risk_Mavaddat2019_cal - (quantile(TC_10risk_Mavaddat2019_cal, 
                                                                 probs=c(1-TC_5prop)) - 0.05),
  TC_Mav2019_10risk_adj_cat = factor(ifelse(TC_10risk_Mavaddat2019_cal <= TC_prs_threshold, 
                                          paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                          paste0(">", round(100*TC_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                        paste0(">", round(100*TC_prs_threshold,2), "%"))),
  gail_prs313_10risk_adj = gail_prs313_10risk - (quantile(gail_prs313_10risk, probs=c(1-gail_5prop)) - 0.05),
  gail_prs313_10risk_adj_cat = factor(ifelse(gail_prs313_10risk <= gail_prs_threshold, 
                                          paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                          paste0(">", round(100*gail_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                        paste0(">", round(100*gail_prs_threshold,2), "%")))
  )

```

```{r, results=FALSE}

# Net reclassification index, for Table 1

# Tyrer-Cuzick
nri_tc <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_prs313_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

nri_tc_mav <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_Mav2019_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

# Gail
nri_gail <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$gail_10risk_cal, p.new=test_data$gail_prs313_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

nri_ci <- function(nri_output, type="overall", dp=3){
  row <- dplyr::case_when(
    type=="overall" ~ 1,
    type=="case" ~ 2,
    type=="control" ~ 3
  )
  est <- nri_output$nri$Estimate[row]
  lci <- nri_output$nri$Lower[row]
  uci <- nri_output$nri$Upper[row]

  return(paste0(pretty_dp(est, dp=dp), " ", pretty_confint(lci, uci, dp=dp)))
}

```

```{r}

dt <- data.frame(model=c("Tyrer-Cuzick", "Model only", "Model with PRS", "Integrated PRS",
                         "Gail", "Model only", "Model with PRS"),
                 AUC=c(NA,
                       AUC_CI(df=test_data, score="TC_10risk_cal", status = "TEU_BrCa_status", dp=2),
                       AUC_CI(df=test_data, score="TC_prs313_10risk", status = "TEU_BrCa_status", dp=2),
                       AUC_CI(df=test_data, score="TC_10risk_Mavaddat2019_cal", status = "TEU_BrCa_status", dp=2),
                       NA,
                       AUC_CI(df=test_data, score="gail_10risk_cal", status = "TEU_BrCa_status", dp=2),
                       AUC_CI(df=test_data, score="gail_prs313_10risk", status = "TEU_BrCa_status", dp=2)),
                 NRI_overall=c(NA, NA, nri_ci(nri_tc, type="overall", dp=3), nri_ci(nri_tc_mav, type="overall", dp=3), 
                               NA, NA, nri_ci(nri_gail, type="overall", dp=3)),
                 NRI_case=c(NA, NA, nri_ci(nri_tc, type="case", dp=3), nri_ci(nri_tc_mav, type="case", dp=3), 
                               NA, NA, nri_ci(nri_gail, type="case", dp=3)),
                 NRI_control=c(NA, NA, nri_ci(nri_tc, type="control", dp=3), nri_ci(nri_tc_mav, type="control", dp=3), 
                               NA, NA, nri_ci(nri_gail, type="control", dp=3)))

ft <- flextable(dt) %>%
  # merge_v(j = ~ model + NRI_overall + NRI_case + NRI_control) %>%
  # separate_header(sep="_") %>%
  set_header_labels(model="Model", AUC="AUC (95% CI)",
                    NRI_overall="Overall", 
                    NRI_case=paste0("Case (N=", nrow(test_data[test_data$TEU_BrCa_time <= 10 &
                                                                 test_data$TEU_BrCa_status==1,]), ")"), 
                    NRI_control=paste0("Control (N=", nrow(test_data[test_data$TEU_BrCa_time > 10,]), ")")) %>%
  add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
  align(align = "center", part = "all", j=c('AUC', 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
  bg(i= ~ is.na(AUC), bg="lightgrey", part="body") %>%
  merge_at(i=1, j=1:5) %>%
  merge_at(i=5, j=1:5) %>%
  fix_border_issues(part="all") %>%
  set_table_properties(layout="autofit") %>%
  footnote(i=1, j=3, 
           value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk"), 
           ref_symbols="a", part="header") %>%
  footnote(i=2, j=3, 
           value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up."), 
           ref_symbols="b", part="header") %>%
  footnote(i=6, j=1, 
           value=as_paragraph("Integrated PRS is the output from the Tyrer-Cuzick model run with the PRS as an input, formatted as a log(OR)"), 
           ref_symbols="c", part="body")

ft

```

### 2023-11-22 Scripts/Data_Management/Scripts_Cleaning/DuckDB.R

@@ -1,0 +1,108 @@
+
+
+
+#' Reads a database of UKB data and returns a dataset containing the requested columns
+#'
+#' The database contains the data with field codes instead of variable names. It has been pre-processed by the UKB-generated R file to apply categorical variable levels and labels. This loads selecteed variables from the database, applies the chosing naming convention and derives requested variables.
+#'
+#' @param col_list The objects defining derived variables you want in your dataset.
+#' @param db The path to the database you want to extract the variables from.
+#' @param name_map The path to the .csv mapping file containing human-readable names for the raw UKB data fields. 
+#'
+#' @return Returns a data.frame containing the variables requested.
+#'
+#' @import DBI
+#' @import duckdb
+#' @export
+#' @examples
+#' \dontrun{
+#' # Extract variables for HTN project from V2 data
+#'
+#' DB_extract(HTNcols, db="ukb_v2.db")
+#' }
+#'
+DB_extract <- function(extract_cols, db = config$data$database, 
+                       name_map = config$cleaning$renaming,
+                       withdrawals = config$cleaning$withdrawals){
+  
+  mapping <- read.csv(name_map, stringsAsFactors = FALSE)
+  withdrawals <- read.csv(withdrawals, header=FALSE)
+  withdrawn_ids <- withdrawals$V1
+  
+  # Connect to the database
+  con <- dbConnect(duckdb::duckdb(), db, read_only=TRUE)
+  on.exit(dbDisconnect(con, shutdown=TRUE))
+  
+  # List all tables available in the database
+  # Each should correspond to one data download
+  tables <- dbListTables(con)
+  tables <- tables[!endsWith(tables, "_stata")]
+  
+  # Join all download tables to get all data and extract requested columns
+  view <- lapply(tables, function(x) tbl(con, from=x)) %>% 
+    reduce(inner_join, by = "f.eid", suffix = c("", ".delete")) %>%
+    select(any_of(name_to_fdot(extract_cols, mapping)), -ends_with(".delete")) %>%
+    filter(!(f.eid %in% withdrawn_ids)) %>% # Exclude participants who have withdrawn
+    filter(f.eid != 6025392) %>% # NOTE: HACKY FIX TO DEAL WITH BROKEN PARTICIPANT
+    collect %>%
+    rename_with(fdot_to_name, mapping=mapping)
+  
+  return(view)
+}
+
+
+
[…]

### 2023-11-22 Scripts/Data_Management/Scripts_Cleaning/TEU_specifications.R

@@ -1,0 +1,155 @@
+# Jennifer Collister
+# 30/09/20
+
+# Load the project config file for filepaths etc
+if (!exists("config")) {
+  library(yaml)
+  config = yaml.load_file("config.yml")
+}
+
+
+specs <- function() {
+  
+  # If you're here to write a new spec, you can run this source line interactively
+  # to load all the variable derivation objects into your working environment
+  # so you get autocomplete when typing them!
+  source(file.path(config$scripts$cleaning, "derivation_objects.R"),
+         local = if (sys.nframe() == 0L) {
+           FALSE
+           } else {
+             TEUmaps <- new.env()
+             }
+         )
+  if (exists("TEUmaps")) {
+    attach(TEUmaps)
+    on.exit(detach(TEUmaps))
+  }
+  
+  # Dataset specifications
+  
+  TEUvars_common <- list(
+    ID,
+    BaC_Sex,
+    TEU_BaC_DateOfBirth,
+    TEU_BaC_YoB,
+    Rec_DateAssess,
+    BaC_Age,
+    BaC_AgeRecruit,
+    TEU_BaC_AgeAtRec,
+    TEU_BaC_AgeCat,
+    TEU_ethnicgrp,
+    TEU_Rec_AssessCentre,
+    TEU_Rec_Country,
+    BSM_BMI,
+    TEU_BSM_BMIcat
+  )
+  
+  UKB_genetic <- list(
+    ID,
+    GeP_UsedInPCA, # Identifies participants which met UKB QC for inclusion in PCA
+    GeP_Outliers, # Identifies participants who are outliers for missingness and heterozygosity
+    GeP_ethnic, # Identifies participants with genetic White British ancestry
+    GeP_Array,
+    GeP_PC(pc=1),
+    GeP_PC(pc=2),
+    GeP_PC(pc=3),
+    GeP_PC(pc=4),
+    GeP_PC(pc=5),
+    GeP_PC(pc=6),
+    GeP_PC(pc=7),
+    GeP_PC(pc=8),
+    GeP_PC(pc=9),
+    GeP_PC(pc=10), # Genetic Principal Components of ancestry
+    GeP_Sex, # Used to check for sex discordance
+    BaC_Sex # Used to check for sex discordance
+  )
+
+  
[…]

### 2023-11-22 Scripts/Data_Management/Scripts_Cleaning/basic_functions.R

@@ -1,0 +1,873 @@
+# Jennifer Collister
+# 30/03/2020
+# Create functions 
+library(data.table)
+library(tools)
+
+# Load the project config file for filepaths etc
+if (!exists("config")) {
+  library(yaml)
+  config = yaml.load_file("config.yml")
+}
+
+read.csv_kdrive <- function(filename, maxErrors=2, sleep=20) {
+  attempts <- 0
+  retval <- NULL
+  
+  while(is.null(retval) && attempts < maxErrors) {
+    if(attempts > 0) {Sys.sleep(sleep)}
+    attempts <- attempts + 1
+    retval <- tryCatch(suppressWarnings(
+      read.csv(filename)),
+      error = function(cond) {
+        message(paste0("Error message: ", cond))
+        if(attempts < maxErrors) {message(paste0("Trying again in ", sleep, " seconds..."))}
+        return(NULL)
+      }
+    )
+  }
+  if(is.null(retval)){
+    message("Failed to read file. Either it does not exist (typo?) or you are currently disconnected from the network drives")
+  } else if (attempts>1) {
+    message("Succeeded!")
+  }
+  
+  return(retval)
+}
+
+read.xlsx_kdrive <- function(filename,col_types=NULL, sheet=NULL, maxErrors=2, sleep=20) {
+  attempts <- 0
+  retval <- NULL
+  
+  while(is.null(retval) && attempts < maxErrors) {
+    if(attempts > 0) {Sys.sleep(sleep)}
+    attempts <- attempts + 1
+    retval <- tryCatch(suppressWarnings(
+      read_excel(filename,col_types = col_types, sheet = sheet)),
+      error = function(cond) {
+        message(paste0("Error message: ", cond))
+        if(attempts < maxErrors) {message(paste0("Trying again in ", sleep, " seconds..."))}
+        return(NULL)
+      }
+    )
+  }
+  if(is.null(retval)){
+    message("Failed to read file. Either it does not exist (typo?) or you are currently disconnected from the network drives")
+  } else if (attempts>1) {
+    message("Succeeded!")
+  }
+  
+  return(retval)
+}
+
+cache.file <- function(filename, FUN, ...) {
+  file <- evalWithMemoization(
+    FUN(filename, ...),
+    key = c(filename, file.info(filename)$mtime)
+  )
+}
[…]

### 2023-11-22 Scripts/Data_Management/Scripts_Cleaning/dataset_generator.R

@@ -1,0 +1,238 @@
+# Jennifer Collister
+# 22/09/20
+
+library(DBI)
+library(duckdb)
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+library(R.cache)
+library(tictoc)
+setCacheRootPath(path="Data/Cache")
+
+# Load the project config file for filepaths etc
+if (!exists("config")) {
+  library(yaml)
+  config = yaml.load_file("config.yml")
+}
+
+# Create an environment for the dependency functions
+DBfunc <- new.env()
+
+# Source the variable maps into the TEUmaps environment
+source(file.path(config$scripts$cleaning, "TEU_specifications.R"))
+source(file.path(config$scripts$cleaning, "DuckDB.R"), local=DBfunc)
+
+# Note - may in future want to use https://cran.r-project.org/web/packages/modules/
+# This means stuff in the modules *can't* see and interact with stuff in global env
+# Better practice
+
+
+derive_variables <- function(database, field_definitions, exclusions=function(x){x}, dictionary=NULL,
+                             name_map = config$cleaning$renaming,
+                             withdrawals = config$cleaning$withdrawals){
+  
+  # Extract the lists from the list of functions
+  objects <- Map(function(p) {if(is.function(p)) {p()} else {p}}, field_definitions)
+  
+  # Find the names of the requested columns
+  outcols <- sapply(objects, function(x) x$name)
+  
+  # Remove duplicated variables that have been entered more than once in spec
+  objects <- objects[!duplicated(outcols)]
+  outcols <- outcols[!duplicated(outcols)]
+  
+  # And the names of all the source columns required
+  source_cols <- unique(unlist(sapply(objects, function(x) x$source)))
+  
+  # Extract data fields from database
+  data <- DBfunc$DB_extract(source_cols, db = database, name_map = name_map, withdrawals = withdrawals)
+
+  # Separate into derivations to be calculated before and after exclusion criteria
+  before <- objects[sapply(objects, function(x) x$post_exclusion==FALSE)]
+  after <- objects[sapply(objects, function(x) x$post_exclusion==TRUE)]
+  
+  data <- DBfunc$source_rotation(data, field_definitions = before)
+  
+  data <- exclusions(data)
+  # Note - future update
+  # Make exclusions a list of functions, each function is a separate exclusion criterion
+  # Then the derive_variables() function can handle the row counts
[…]

### 2023-11-22 Scripts/Data_Management/Scripts_Cleaning/derivation_objects.R

@@ -1,0 +1,2107 @@
+# Jennifer Collister
+# 22/09/20
+
+library(glue)
+library(lubridate)
+library(readxl)
+
+# Load the project config file for filepaths etc
+if (!exists("config")) {
+  library(yaml)
+  config = yaml.load_file("config.yml")
+}
+# Check the 
+showcase_censoring <- yaml.load_file(file.path(dirname(config$data$database), "censoring.yml"))
+cancercol_icd9 <- showcase_censoring$Cancer$icd9_col
+cancercol_icd10 <- showcase_censoring$Cancer$icd10_col
+
+# Source the function definitions
+source(file.path(config$scripts$cleaning, "basic_functions.R"), local = TRUE)
+
+
+# Formatting of existing UKB variables
+
+ID <- function() {
+  list(
+    name = "ID",
+    source = "ID",
+    mapper = FN_id,
+    post_exclusion = FALSE,
+    display_name = "ID",
+    description = "The unique participant identifier"
+  )
+}
+
+BaC_Sex <- function() {
+  list(
+    name = "BaC_Sex",
+    source = "BaC_Sex.0.0",
+    mapper = FN_unorder,
+    post_exclusion = FALSE,
+    display_name = "Gender",
+    description = "Participant's self-reported gender"
+  )
+}
+
+Rec_DateAssess <- function() {
+  list(
+    name = "Rec_DateAssess",
+    source = c("Rec_DateAssess.0.0"),
+    mapper = FN_toDate,
+    post_exclusion = FALSE,
+    display_name = "Date of baseline assessment",
+    description = "Date of baseline assessment"
+  )
+}
+
+Eth_Ethnicity <- function() {
+  list(
+    name = "Eth_Ethnicity",
+    source = "Eth_Ethnicity.0.0",
+    mapper = FN_factor(
+      levelorder = c(
+        "White",
+        "British",
+        "Irish",
+        "Any other white background",
+        "Mixed",
+        "White and Black Caribbean",
+        "White and Black African",
+        "White and Asian",
+        "Any other mixed background",
+        "Asian or Asian British",
+        "Indian",
+        "Pakistani",
+        "Bangladeshi",
+        "Any other Asian background",
+        "Black or Black British",
+        "Caribbean",
+        "African",
+        "Any other Black background",
+        "Chinese",
+        "Other ethnic group",
[…]

### 2023-11-22 Scripts/Stats_Analysis/JCfunctions.R

@@ -1,0 +1,482 @@
+library(stringr)
+options("scipen"=100)
+
+# Run chunks of R code in arbitrary working directory
+# From Hadley Wickham
+# https://github.com/yihui/knitr/issues/38
+in_dir <- function(dir, code) {
+  cur <- getwd()
+  setwd(dir)
+  on.exit(setwd(cur))
+  
+  force(code)
+}
+
+pretty_dp <- function(x, dp=0, pct=FALSE, comma=FALSE){
+  if(pct){x <- 100*x}
+  if(comma){
+    format(round(x, dp), digits=dp, nsmall=dp, big.mark=",") %>% trimws
+  } else {
+    format(round(x, dp), digits=dp, nsmall=dp) %>% trimws
+  }
+}
+
+pretty_confint <- function(lci, uci, dp, pct=FALSE){
+  paste0("(", pretty_dp(x=lci, dp=dp, pct=pct), ", ", pretty_dp(x=uci, dp=dp, pct=pct), ")")
+}
+
+pretty_pval <- function(p, cutoff=0.001, string="<0.001", dp=3){
+  ifelse(p<cutoff, string, pretty_dp(p, dp))
+}
+
+lower <- function(x){
+  paste0(tolower(substring(x, 1,1)), substring(x, 2))
+}
+
+upper <- function(x){
+  paste0(toupper(substring(x, 1,1)), substring(x, 2))
+}
+
+prettyfunc <- function(x, pnames=list(), upper=FALSE, bold=FALSE, flist=c()){
+  out <- x
+  if(x %in% names(pnames)){
+    out <- pnames[[x]]
+    if(upper){
+      out <- upper(out)
+    }
+    if(bold){
+      out <- paste0("**", out, "**")
+    }
+  }
+  if(x %in% flist){
+    if(!exists("footnote_no")){
+      footnote_no <<- 1
+    }
+    out <- paste0(out, "^", footnote_no, "^")
+    footnote_no <<- footnote_no + 1
+  }
+  return(out)
+}
+
+diagcollist <- function(colstring, sep="", ncols) {
+  x <- 0:ncols
+  colstr <- paste0("`",paste0(colstring, sep, x, collapse="`, `"),"`")
+  return(colstr)
+}
+
+# Print numbers and proportions for factors, median and IQR or mean and 95% CI for continuous variables
+# Optionally provide p-values from chi-squared (categorical) and t-test (continuous)
[…]

### 2024-01-23 Scripts/Data_Management/Scripts_Cleaning/basic_functions.R
Database connections should be read-only

@@ -461,7 +461,7 @@ FN_HES_diagnoses <- function(icd=10,
   
   # Connect to database
   dbname <- file.path(HESPath, "hes.db")
-  con <- DBI::dbConnect(duckdb::duckdb(), dbname)
+  con <- DBI::dbConnect(duckdb::duckdb(), dbname, read_only = TRUE)
   on.exit(DBI::dbDisconnect(con, shutdown=TRUE))
   
   hesin <- tbl(con, "hesin")
@@ -506,7 +506,7 @@ FN_HES_operations <- function(opcs=4,
   
   # Connect to database
   dbname <- file.path(HESPath, "hes.db")
-  con <- DBI::dbConnect(duckdb::duckdb(), dbname)
+  con <- DBI::dbConnect(duckdb::duckdb(), dbname, read_only = TRUE)
   on.exit(DBI::dbDisconnect(con, shutdown=TRUE))
   
   hesin <- tbl(con, "hesin")

### 2024-01-24 Scripts/Data_Management/Scripts_Processing/Prep_TyrerCuzick_input.Rmd
R1.7 no FaH BC father

@@
   v8 <- data %>% transmute(
-    t1_FatherBrCa = dplyr::case_when(
-      TC_FatherBrCa=="Yes" ~ 1,
-      TC_FatherBrCa=="No" ~ 0,
-      ), # All No but include anyway for completion
+    t1_FatherBrCa = 0, # No data
     t2_BrotherBrCa = dplyr::case_when(

### 2024-01-24 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
R1.7 no FaH BC father

@@
                    TC_MotherBrCa + 
-                   TC_NSisters + TC_SiblingBrCa + TC_FatherBrCa,
+                   TC_NSisters + TC_SiblingBrCa,
                  data=data %>% mutate(Biopsy_Prevalent=factor(Biopsy_Prevalent)), 
@@
                 # "AtypicalHyperplasia", "TC_OvarianCaDx", "TC_OvarianCaAge", "FSF_HRTDuration", 
-                # "FSF_HRT_timesince", "TC_MotherBrCa", "TC_NSisters", "TC_SiblingBrCa", "TC_FatherBrCa")
+                # "FSF_HRT_timesince", "TC_MotherBrCa", "TC_NSisters", "TC_SiblingBrCa")
 

### 2024-01-24 Scripts/Data_Management/Scripts_Cleaning/derivation_objects.R

@@ -1881,19 +1881,6 @@ TC_FaH <- function(){
       post_exclusion = FALSE,
       display_name = "Sibling(s) breast cancer status",
       description = "Has a sibling had breast cancer, set to 0 if not relevant, missing or unknown"
-    ),
-    list(
-      name = "TC_FatherBrCa",
-      source = c("ID", paste0("FaH_FatherIll.0.", c(0:9))),
-      mapper = function(data){
-        brca <- data %>% pivot_longer(cols=starts_with("FaH_FatherIll.0.")) %>%
-          filter(value =="Breast cancer")
-        y <- ifelse(data[["ID"]] %in% brca$ID, "Yes", "No")
-        return(y)
-      },
-      post_exclusion = FALSE,
-      display_name = "Paternal breast cancer status",
-      description = "Has the father had breast cancer, set to 0 if not relevant, missing or unknown"
     )
   )
   

### 2024-01-24 Scripts/Data_Management/Scripts_Processing/Prep_TyrerCuzick_input.Rmd
R1.3 adjust genomics PRS

@@
 #-------------#
-# Run 3 times #
+# Run 4 times #
 #-------------#
 # t5_PRS = 0, filename "noPRS"
+# t5_PRS = PRS_bc_adj, filename "withPRS/GenomicsPLC"
 # t5_PRS = TEU_BrCa_313_PRS_center, filename "withPRS/Mavaddat2019"

### 2024-01-24 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
R1.3 adjust genomics PRS

@@
 
+# TC v8 with Genomics plc
+
+```{r}
+
+TC_5pct <- table(test_data$TC_10risk_cal > 0.05, useNA='ifany')
+TC_5prop <- TC_5pct[2]/nrow(test_data)
+TC_Gplc_threshold <- quantile(test_data$TC_10risk_GenomicsPLC_cal, probs=c(1-TC_5prop))
+
+test_data <- test_data %>% mutate(
+  TC_Gplc_10risk_adj = TC_10risk_GenomicsPLC_cal - (quantile(TC_10risk_GenomicsPLC_cal, 
+                                                                 probs=c(1-TC_5prop)) - 0.05),
+  TC_Gplc_10risk_adj_cat = factor(ifelse(TC_10risk_GenomicsPLC_cal <= TC_Gplc_threshold, 
+                                          paste0("<=", round(100*TC_Gplc_threshold,2), "%"), 
+                                          paste0(">", round(100*TC_Gplc_threshold,2), "%")),
+                               levels=c(paste0("<=", round(100*TC_Gplc_threshold,2), "%"), 
+                                        paste0(">", round(100*TC_Gplc_threshold,2), "%")))
+  )
+
+nri_tc_Gplc <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
+        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_Gplc_10risk_adj,
+        t0=10, cut=0.05, msg=FALSE)
+
+```
+
+```{r}
+
+dt <- data.frame(model=c("Model only", "Model with PRS", "Integrated PRS"),
+                 HarrellC=c(
+                       HarrellC_CI(df=test_data, score="TC_10risk_cal", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="TC_prs_10risk", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="TC_10risk_GenomicsPLC_cal", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2)),
+                 NRI_overall=c(NA, nri_ci(nri_tc, type="overall", dp=3), 
+                               nri_ci(nri_tc_Gplc, type="overall", dp=3)),
+                 NRI_case=c(NA, nri_ci(nri_tc, type="case", dp=3), 
+                            nri_ci(nri_tc_Gplc, type="case", dp=3)),
+                 NRI_control=c(NA, nri_ci(nri_tc, type="control", dp=3), 
+                               nri_ci(nri_tc_Gplc, type="control", dp=3)))
+
+ft <- flextable(dt) %>%
+  # merge_v(j = ~ model + NRI_overall + NRI_case + NRI_control) %>%
+  # separate_header(sep="_") %>%
+  set_header_labels(model="Tyrer-Cuzick Model", HarrellC="Harrell's C (95% CI)",
+                    NRI_overall="Overall", 
+                    NRI_case=paste0("Case (N=", nrow(test_data[test_data$TEU_BrCa_time <= 10 &
+                                                                 test_data$TEU_BrCa_status==1,]), ")"), 
+                    NRI_control=paste0("Control (N=", nrow(test_data[test_data$TEU_BrCa_time > 10,]), ")")) %>%
+  add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
+  align(align = "center", part = "all", j=c("HarrellC", 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
+  bg(i= ~ is.na(HarrellC), bg="lightgrey", part="body") %>%
+  fix_border_issues(part="all") %>%
+  set_table_properties(layout="autofit") %>%
+  flextable::footnote(i=1, j=3, ref_symbols="a", part="header",
+           value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk")) %>%
+  flextable::footnote(i=2, j=3, ref_symbols="b", part="header",
+           value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up.")) %>%
+  flextable::footnote(i=3, j=1, ref_symbols="c", part="body",
+           value=as_paragraph("Integrated PRS is the output from the Tyrer-Cuzick model run with the PRS as an input, formatted as a log(OR)"))
+
+ft
+
+```
+
 # Results using Mavaddat 313-SNP PRS

### 2024-01-24 Scripts/Data_Management/Scripts_Cleaning/TEU_specifications.R

@@ -143,8 +143,10 @@ specs <- function() {
     list(
       train_set,
       TC_risk(dir=file.path(config$data$derived), years=10, prs=NULL),
+      TC_risk(dir=file.path(config$data$derived), years=10, prs="GenomicsPLC"),
       TC_risk(dir=file.path(config$data$derived), years=10, prs="Mavaddat2019"),
       TC_risk_cal(years=10, prs=NULL),
+      TC_risk_cal(years=10, prs="GenomicsPLC"),
       TC_risk_cal(years=10, prs="Mavaddat2019")
     )
   )

### 2024-01-24 Scripts/Data_Management/Scripts_Cleaning/derivation_objects.R

@@ -1519,6 +1519,22 @@ genomics_prs <- function(){
       description = "Genomics plc standard PRS for breast cancer"
     ),
     list(
+      name = "PRS_bc_adj",
+      source = c("PRS_bc_std"),
+      mapper = function(x) {
+        PRS_std <- (x - mean(x)) / sd(x)
+        
+        # log(OR) per SD of PRS, from model with PRS only, from supplementary Materials of Thompson2022 (S7 Performance (diseases))
+        logORperSD <- log(1.82383417)
+        
+        PRS_adj <- PRS_std - ((logORperSD^2)/2)
+        return(PRS_adj)
+      },
+      post_exclusion = TRUE,
+      display_name = "Genomics plc standard PRS for breast cancer",
+      description = "Genomics plc standard PRS for breast cancer, adjusted such that the mean predicted risk is equal to the population mean risk as per Pharoah et al 2002"
+    ),
+    list(
       name = "PRS_bc_std_q",
       source = c("PRS_bc_std"),
       mapper = FN_quantiles(

### 2024-01-24 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Swap AUC to Harrell's C

--- 2023-11-22 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
@@
                          "Gail", "Model only", "Model with PRS"),
-                 AUC=c(NA,
-                       AUC_CI(df=test_data, score="TC_10risk_cal", status = "TEU_BrCa_status", dp=2),
-                       AUC_CI(df=test_data, score="TC_prs_10risk", status = "TEU_BrCa_status", dp=2),
+                 HarrellC=c(NA,
+                       HarrellC_CI(df=test_data, score="TC_10risk_cal", 
+                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="TC_prs_10risk", 
+                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                        NA,
-                       AUC_CI(df=test_data, score="gail_10risk_cal", status = "TEU_BrCa_status", dp=2),
-                       AUC_CI(df=test_data, score="gail_prs_10risk", status = "TEU_BrCa_status", dp=2)),
+                       HarrellC_CI(df=test_data, score="gail_10risk_cal", 
+                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="gail_prs_10risk", 
+                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2)),
                  NRI_overall=c(NA, nri_ci(nri_tc, type="overall", dp=3), nri_ci(nri_tc, type="overall", dp=3), 
@@
   # separate_header(sep="_") %>%
-  set_header_labels(model="Model", AUC="AUC (95% CI)",
+  set_header_labels(model="Model", HarrellC="Harrell's C (95% CI)",
                     NRI_overall="Overall", 
@@
   add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
-  align(align = "center", part = "all", j=c('AUC', 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
-  bg(i= ~ is.na(AUC), bg="lightgrey", part="body") %>%
+  align(align = "center", part = "all", j=c("HarrellC", 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
+  bg(i= ~ is.na(HarrellC), bg="lightgrey", part="body") %>%
   merge_at(i=1, j=1:5) %>%
@@
                          "Gail", "Model only", "Model with PRS"),
-                 AUC=c(NA,
-                       AUC_CI(df=test_data, score="TC_10risk_cal", status = "TEU_BrCa_status", dp=2),
-                       AUC_CI(df=test_data, score="TC_prs313_10risk", status = "TEU_BrCa_status", dp=2),
-                       AUC_CI(df=test_data, score="TC_10risk_Mavaddat2019_cal", status = "TEU_BrCa_status", dp=2),
+                 HarrellC=c(NA,
+                       HarrellC_CI(df=test_data, score="TC_10risk_cal", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="TC_prs313_10risk", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="TC_10risk_Mavaddat2019_cal", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                        NA,
-                       AUC_CI(df=test_data, score="gail_10risk_cal", status = "TEU_BrCa_status", dp=2),
-                       AUC_CI(df=test_data, score="gail_prs313_10risk", status = "TEU_BrCa_status", dp=2)),
+                       HarrellC_CI(df=test_data, score="gail_10risk_cal", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
+                       HarrellC_CI(df=test_data, score="gail_prs313_10risk", 
+                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2)),
                  NRI_overall=c(NA, NA, nri_ci(nri_tc, type="overall", dp=3), nri_ci(nri_tc_mav, type="overall", dp=3), 

### 2024-01-24 Scripts/Stats_Analysis/JCfunctions.R

@@ -480,3 +480,17 @@ AUC_CI<-function(df,status="TEU_T2DM_status",score,method="delong", dp=3){
 }
 
 
+HarrellC_CI <- function(df, status, time, score, dp=3) {
+  # https://www.rdocumentation.org/packages/Hmisc/versions/5.1-1/topics/rcorr.cens
+  
+  stat <- Hmisc::rcorr.cens(x=df[[score]], S=Surv(time=df[[time]], event=df[[status]]))
+  
+  HarrellC <- 1 - stat["C Index"]
+  se <- stat["S.D."]/2
+  LCI <- HarrellC - (1.96 * se)
+  UCI <- HarrellC + (1.96 * se)
+  
+  result <- paste0(pretty_dp(HarrellC, dp), pretty_confint(LCI, UCI, dp=dp))
+  
+  return(result)
+}

### 2024-01-24 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
ROC plots are now time-dependent

@@
 library(maps)
-library(pROC)
+library(timeROC)
 
@@
 
-# ROC plot
-# https://rviews.rstudio.com/2019/03/01/some-r-packages-for-roc-curves/
-# https://cran.r-project.org/web/packages/plotROC/plotROC.pdf
-# https://cran.r-project.org/web/packages/plotROC/vignettes/examples.html
-
-# New approach uses pROC, Xiaonan's code
-# https://gitlab.ndph.ox.ac.uk/jenniferco/prsondiabetes/-/blob/Riskscore/Scripts/Stats_Analysis/RMarkdown/RiskScore_Primary.Rmd#auc
-
-roc_plot <- function(data, modelname, prsname, plotname,
+# Time-dependent ROC at 10 years, using timeROC
+
+td_roc_plot <- function(data, modelname, prsname, plotname,
                      varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk")){
 
-  AUC_CIs <- sapply(varlist, function(x) AUC_CI(df=data, score=x, status = "TEU_BrCa_status", dp=2))
-
-  AUC_labels <- c(
-    paste0(modelname, " 10-yr risk: ", AUC_CIs[1]),
-    paste0(prsname, " PRS: ", AUC_CIs[2]),
-    paste0(modelname, " & PRS 10-yr risk: ", AUC_CIs[3])
-  )
-
-  long <- data %>%
-    select(ID, TEU_BrCa_status, all_of(varlist)) %>%
-    pivot_longer(!c("ID","TEU_BrCa_status"), names_to = "Score", values_to = "Value") %>%
-    mutate(Score=factor(Score,levels = varlist))
-
-  plot <- ggplot(long, aes(d=TEU_BrCa_status, m=Value, color=Score)) +
-    geom_roc(n.cuts = 0, linealpha = 1, size=1) +
-    style_roc() +
+  # Time-dependent ROC
+  tROC <- lapply(varlist,
+                 function(i) timeROC::timeROC(T=data$TEU_BrCa_time, 
+                                              delta=data$TEU_BrCa_status, 
+                                              marker=data[[i]],
+                                              cause=1, times=c(9, 10), weighting="marginal"))
+
+  tROC_long <- rbind(cbind(tROC[[1]][["TP"]], tROC[[1]][["FP"]]),
+                     cbind(tROC[[2]][["TP"]], tROC[[2]][["FP"]]),
+                     cbind(tROC[[3]][["TP"]], tROC[[3]][["FP"]]))
+  colnames(tROC_long) <- c("TP_9", "TP_10", "FP_9", "FP_10")
+  tROC_df <- data.frame(tROC_long)
+  tROC_df$Model <- factor(c(rep(paste0(modelname, " 10-yr risk"), times=dim(tROC[[1]][["TP"]])[1]),
+                     rep(paste0(prsname, " PRS"), times=dim(tROC[[2]][["TP"]])[1]),
+                     rep(paste0(modelname, " & PRS 10-yr risk"), times=dim(tROC[[3]][["TP"]])[1])),
+                     levels=c(paste0(modelname, " 10-yr risk"),
+                              paste0(prsname, " PRS"),
+                              paste0(modelname, " & PRS 10-yr risk")))
+
+  plot <- ggplot(tROC_df,aes(x=FP_10, y=TP_10, color=Model))+
+    geom_line(linewidth=1) + style_roc() +
     geom_abline(aes(slope=1, intercept=0), linetype="dashed") +
-    scale_color_manual(name="AUC (95% CI)", labels =  AUC_labels, values = c("#F8766D","#00BA38", "#6193FF")) +
-    labs(title = paste0(plotname), 
-         caption = "Dashed line is y=x") +
-    theme(legend.position = c(0.7, 0.15))
+    xlab("False positive fraction") + ylab("True positive fraction") +
+    scale_color_manual(name="Model",values = c("#F8766D","#00BA38", "#6193FF")) +
+    labs(title =plotname, caption = "Dashed line is y=x")+
+    theme(legend.position = c(0.75, 0.15))
 
@@
 # Tyrer-Cuzick
-TC_roc <- roc_plot(test_data, modelname="TC", prsname="Genomics plc", plotname="Tyrer-Cuzick",
+TC_roc <- td_roc_plot(test_data, modelname="TC", prsname="Genomics plc", plotname="Tyrer-Cuzick",
          varlist=c("TC_10risk_cal", "PRS_bc_std", "TC_prs_10risk"))
@@
 # Gail
-gail_roc <- roc_plot(test_data, modelname="Gail", prsname="Genomics plc", plotname="Gail",
+gail_roc <- td_roc_plot(test_data, modelname="Gail", prsname="Genomics plc", plotname="Gail",
          varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk"))
@@
                         ncol=2,
-                        top=paste0("ROC curves for 10-yr risk predictions in test data (N=", nrow(test_data), ")"))
+                        top=paste0("Time-dependent ROC curves at 10 years follow up in test data (N=", nrow(test_data), ")"))
 

### 2024-01-24 Scripts/Stats_Analysis/JCfunctions.R

@@ -467,18 +467,6 @@ cuminc_nocr_plot<-function(data,start,end,status,group,title,legend.title,xlab="
   return(plot)
 }
 
-AUC_CI<-function(df,status="TEU_T2DM_status",score,method="delong", dp=3){
-  
-  roccurve<-pROC::roc(df[[status]]~df[[score]],quiet=TRUE)
-  
-  AUCs_train<-pROC::ci.auc(roccurve,quiet=TRUE,method=method)
-  
-  result<-paste0(pretty_dp(AUCs_train[2],dp), " ", pretty_confint(AUCs_train[1],AUCs_train[3],dp=dp))
-  
-  return(result)
-  
-}
-
 
 HarrellC_CI <- function(df, status, time, score, dp=3) {
   # https://www.rdocumentation.org/packages/Hmisc/versions/5.1-1/topics/rcorr.cens

### 2024-01-24 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Assorted improvements and bugfixes

--- 2024-01-24 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
@@
   set_table_properties(layout="autofit") %>%
-  footnote(i=1, j=1, 
+  flextable::footnote(i=1, j=1, 
            value=as_paragraph("Hazard ratio per standard deviation from a Cox model containing 10-year risk score from breast cancer risk prediction model, Genomics plc PRS, genetic array and first 4 principal components of genetic ancestry, in the training data"),
@@
   set_table_properties(layout="autofit") %>%
-  footnote(i=1, j=3, 
+  flextable::footnote(i=1, j=3, 
            value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk"), 
            ref_symbols="a", part="header") %>%
-  footnote(i=2, j=3, 
+  flextable::footnote(i=2, j=3, 
            value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up."),
@@
     fix_border_issues(part="all") %>%
-    footnote(i=1, j=1, value=as_paragraph("Cases defined as individuals diagnosed with breast cancer within 10 years.
+    flextable::footnote(i=1, j=1, value=as_paragraph("Cases defined as individuals diagnosed with breast cancer within 10 years.
 Controls defined as individuals who were still at risk of breast cancer by 10 years of follow-up.
@@
   select(ID, BaC_Age, TEU_BrCa_time, TEU_BrCa_status, 
-                                 tc_varlist) %>%
+                                 all_of(tc_varlist)) %>%
   drop_na
@@
   select(ID, BaC_Age, TEU_BrCa_time, TEU_BrCa_status, 
-                                 gail_varlist) %>%
+                                 all_of(gail_varlist)) %>%
   drop_na
@@
   TC_Mav2019_10risk_adj_cat = factor(ifelse(TC_10risk_Mavaddat2019_cal <= TC_prs_threshold, 
-                                          paste0("<=", round(100*TC_prs_threshold,2), "%"), 
-                                          paste0(">", round(100*TC_prs_threshold,2), "%")),
-                               levels=c(paste0("<=", round(100*TC_prs_threshold,2), "%"), 
-                                        paste0(">", round(100*TC_prs_threshold,2), "%"))),
+                                          paste0("<=", round(100*TC_Mav2019_threshold,2), "%"), 
+                                          paste0(">", round(100*TC_Mav2019_threshold,2), "%")),
+                               levels=c(paste0("<=", round(100*TC_Mav2019_threshold,2), "%"), 
+                                        paste0(">", round(100*TC_Mav2019_threshold,2), "%"))),
   gail_prs313_10risk_adj = gail_prs313_10risk - (quantile(gail_prs313_10risk, probs=c(1-gail_5prop)) - 0.05),
@@
   set_table_properties(layout="autofit") %>%
-  footnote(i=1, j=3, 
+  flextable::footnote(i=1, j=3, 
            value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk"), 
            ref_symbols="a", part="header") %>%
-  footnote(i=2, j=3, 
+  flextable::footnote(i=2, j=3, 
            value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up."),
            ref_symbols="b", part="header") %>%
-  footnote(i=6, j=1, 
+  flextable::footnote(i=4, j=1, 
            value=as_paragraph("Integrated PRS is the output from the Tyrer-Cuzick model run with the PRS as an input, formatted as a log(OR)"), 

### 2024-01-24 Scripts/Data_Management/Scripts_Cleaning/dataset_generator.R

@@ -143,6 +143,7 @@ DBfunc$source_rotation <- function(data, field_definitions) {
 
 # Actually do the deriving
 DBfunc$derive_fn <- function(data, field_definition) {
+  print(paste0("Deriving ", field_definition$name))
   tic()
   colfunc <- field_definition$mapper
   if(length(field_definition$source)>1){
@@ -150,7 +151,6 @@ DBfunc$derive_fn <- function(data, field_definition) {
   } else {
     data[[field_definition$name]] <- colfunc(data[[field_definition$source]])
   }
-  print(paste0("Derived ", field_definition$name))
   toc()
 
   return(data)

### 2024-01-25 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Calibrate using 80% training

@@
 data <- data %>% mutate(
-  TC_10risk_sd = TC_10risk_cal/sd(TC_10risk_cal),
-  gail_10risk_sd = gail_10risk_cal/sd(gail_10risk_cal),
+  TC_10risk_sd = TC_10risk/sd(TC_10risk),
+  gail_10risk_sd = gail_10risk/sd(gail_10risk),
   PRS_bc_sd = PRS_bc_std/sd(PRS_bc_std),
@@
 
-cal_plot <- function(data, var="TC_10risk", title, xlim=c(0.01,0.07), ylim=c(0.01,0.07)){
+cal_plot <- function(data, var="TC_10risk", title, xlim=c(0, 0.1), ylim=c(0, 0.1)){
 
@@
 
-  return(plot)
+  output <- list(plot=plot, cal_slope=cal_slope, cal_intercept=cal_intercept)
+
+  return(output)
 
@@
 
-TC_cal <- cal_plot(data=data, var="TC_10risk",
-                   title="Tyrer-Cuzick model")
-gail_cal <- cal_plot(data=data, var="gail_10risk",
-                     title="Gail model") +
-  theme(legend.position="none")
-TC_cal2 <- cal_plot(data=data, var="TC_10risk_cal",
-                   title="Calibrated Tyrer-Cuzick model")
-gail_cal2 <- cal_plot(data=data, var="gail_10risk_cal",
-                     title="Calibrated Gail model") +
-  theme(legend.position="none")
+TC_train <- cal_plot(data=train_data, var="TC_10risk",
+                   title="Tyrer-Cuzick model in training data")
+gail_train <- cal_plot(data=train_data, var="gail_10risk",
+                     title="Gail model in training data")
+
+TC_test <- cal_plot(data=test_data, var="TC_10risk_cal",
+                   title="Calibrated Tyrer-Cuzick model in test data")
+gail_test <- cal_plot(data=test_data, var="gail_10risk_cal",
+                     title="Calibrated Gail model in test data")
+
+TC_prs_cal <- cal_plot(data=test_data, var="TC_prs_10risk",
+                   title="Tyrer-Cuzick model with PRS in test data")
+gail_prs_cal <- cal_plot(data=test_data, var="gail_prs_10risk",
+                     title="Gail model with PRS in test data")
 
@@
 #+++++++++++++++++++++++
-legend <- get_legend(TC_cal)
-# 3. Remove the legend from the box plot
+legend <- get_legend(TC_train$plot)
+# 3. Remove the legend
 #+++++++++++++++++++++++
-TC_cal <- TC_cal + theme(legend.position="none")
-
-gridExtra::grid.arrange(TC_cal, gail_cal, legend, TC_cal2, gail_cal2,
-                        ncol=3, widths=c(3.5, 3.5, 0.8),
-                        top=paste0("Before and after calibration plots of 10-year observed vs predicted risk by predicted risk decile of each model in the whole analysis population (N=", nrow(data), ")"))
+gridExtra::grid.arrange(TC_train$plot + theme(legend.position="none"), 
+                        gail_train$plot + theme(legend.position="none"), 
+                        legend, 
+                        TC_test$plot + theme(legend.position="none"), 
+                        gail_test$plot + theme(legend.position="none"),
+                        NULL,
+                        TC_prs_cal$plot + theme(legend.position="none"),
+                        gail_prs_cal$plot + theme(legend.position="none"),
+                        ncol=3, widths=c(3.5, 3.5, 0.8))
 

### 2024-01-25 Scripts/Data_Management/Scripts_Cleaning/derivation_objects.R

@@ -1937,32 +1937,37 @@ TC_risk <- function(dir=config$data$derived, years=10, prs=NULL){
 TC_risk_cal <- function(years=10, prs=NULL) {
   list(
     name = ifelse(!is.null(prs), glue("TC_{years}risk_{prs}_cal"), glue("TC_{years}risk_cal")),
-    source = c(ifelse(!is.null(prs), glue("TC_{years}risk_{prs}"), glue("TC_{years}risk")), 
+    source = c("ID", "train_set", 
+               ifelse(!is.null(prs), glue("TC_{years}risk_{prs}"), glue("TC_{years}risk")), 
                "TEU_BrCa_time", "TEU_BrCa_status"),
     mapper = function(data){
       risk_col <- ifelse(!is.null(prs), glue("TC_{years}risk_{prs}"), glue("TC_{years}risk"))
       
+      train_IDs <- read.csv(file.path(config$data$derived, "training_ids.csv"))$x
+      
       # Calibrate on 10 years of follow-up only
-      data <- data[!is.na(risk_col),] %>% 
+      train_data <- data[!is.na(risk_col),] %>% 
+        # filter(train_set == TRUE) %>%
+        filter(ID %in% train_IDs) %>%
         mutate(
           TEU_BrCa_status = ifelse(TEU_BrCa_time < 10, TEU_BrCa_status, 0),
           TEU_BrCa_time = ifelse(TEU_BrCa_time < 10, TEU_BrCa_time, 10)
           )
       
-      data$original <- data[[risk_col]]
+      train_data$original <- train_data[[risk_col]]
       
-      deciles <- quantile(data$original, probs = seq(0, 1, 0.1))
+      deciles <- quantile(train_data$original, probs = seq(0, 1, 0.1))
       
-      data$risk_dec <- cut(data$original, breaks=deciles, 
+      train_data$risk_dec <- cut(train_data$original, breaks=deciles, 
                               right=TRUE, include.lowest=TRUE) 
       
       # Mean predicted risk per decile
-      predicted <- data %>%
+      predicted <- train_data %>%
         group_by(risk_dec) %>% 
         summarise(predicted = mean(original))
       
       # Fit survival curves from Kaplan-Meier and extract observed risks by predicted risk decile
-      fit <- survfit(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ risk_dec, data=data)
[…]

### 2024-02-10 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Tidy figures, save in format required for resubmission

@@
 
-export_svg(DiagrammeR::grViz(glue(gv))
-           ) %>% charToRaw %>% rsvg %>%
-  png::writePNG(here::here(file.path(config$outputs$figures, "Exclusions.png")))
+# Supplementary Figure 1
+
+export_svg(DiagrammeR::grViz(glue(gv))) %>% 
+  charToRaw %>% 
+  rsvg %>%
+  png::writePNG(here::here(file.path(config$outputs$figures$final, "SuppFigureS1.png")))
+
 ```
@@
 test.ph$y <- y
+
+# Supplementary Figure 3
+tiff(file.path(config$outputs$figures$final,"SuppFigureS3.tiff"), units="in", width=8, height=5, res=300)
 ggcoxzph(test.ph, point.alpha=0.04, var=c("Tyrer-Cuzick 10yr risk", "PRS_BC"),
          caption="Model also includes genetic array and first 4 principal components of genetic ancestry, plots omitted.")
+dev.off()
 
@@
 test.ph$y <- y
+
+# Supplementary Figure 4
+tiff(file.path(config$outputs$figures$final,"SuppFigureS4.tiff"), units="in", width=8, height=5, res=300)
 ggcoxzph(test.ph, point.alpha=0.04, var=c("Gail 10yr risk", "PRS_BC"),
          caption="Model also includes genetic array and first 4 principal components of genetic ancestry, plots omitted.")
+dev.off()
 
@@
 
-cal_plot <- function(data, var="TC_10risk", title, xlim=c(0, 0.1), ylim=c(0, 0.1)){
+cal_plot <- function(data, var="TC_10risk", title, xlim=c(0,0.1), ylim=c(0,0.1)){
 
@@
          title=title) +
-    annotate("text", x=0.05, y = 0.02, size=10/.pt,
+    annotate("text", x=0.07, y = 0.015, size=10/.pt,
              label=paste0("Slope = ", pretty_dp(cal_slope, 3), 
@@
 
-  ggsave(file.path(config$outputs$figures, paste0(var, "_calibration_plot10.jpg")), 
+  ggsave(file.path(config$outputs$figures$interim, paste0(var, "_calibration_plot10.jpg")), 
        scale=0.6, width=25, height=15, units="cm")
@@
 #+++++++++++++++++++++++
+# Figure 1
+tiff(file.path(config$outputs$figures$final,"Figure1.tiff"), units="in", width=8, height=10, res=300)
 gridExtra::grid.arrange(TC_train$plot + theme(legend.position="none"), 
@@
                         gail_prs_cal$plot + theme(legend.position="none"),
-                        ncol=3, widths=c(3.5, 3.5, 0.8))
+                        ncol=3, widths=c(3.5, 3.5, 1))
+dev.off()
 
@@
   tROC_df$Model <- factor(c(rep(paste0(modelname, " 10-yr risk"), times=dim(tROC[[1]][["TP"]])[1]),
-                     rep(paste0(prsname, " PRS"), times=dim(tROC[[2]][["TP"]])[1]),
-                     rep(paste0(modelname, " & PRS 10-yr risk"), times=dim(tROC[[3]][["TP"]])[1])),
+                     rep(prsname, times=dim(tROC[[2]][["TP"]])[1]),
+                     rep(paste0(modelname, " & PRS_BC 10-yr risk"), times=dim(tROC[[3]][["TP"]])[1])),
                      levels=c(paste0(modelname, " 10-yr risk"),
-                              paste0(prsname, " PRS"),
-                              paste0(modelname, " & PRS 10-yr risk")))
+                              prsname,
+                              paste0(modelname, " & PRS_BC 10-yr risk")))
 
@@
     labs(title =plotname, caption = "Dashed line is y=x")+
-    theme(legend.position = c(0.75, 0.15))
-
-  ggsave(file.path(config$outputs$figures, paste0("rocplot_", modelname, ".jpg")),
+    theme(legend.position = c(0.7, 0.15))
+
+  ggsave(file.path(config$outputs$figures$interim, paste0("rocplot_", modelname, ".jpg")),
        scale=0.7, width=25, height=15, units="cm")
@@
 # Tyrer-Cuzick
-TC_roc <- td_roc_plot(test_data, modelname="TC", prsname="Genomics plc", plotname="Tyrer-Cuzick",
+TC_roc <- td_roc_plot(test_data, modelname="Tyrer-Cuzick", prsname="PRS_BC", plotname="Tyrer-Cuzick",
          varlist=c("TC_10risk_cal", "PRS_bc_std", "TC_prs_10risk"))
@@
 # Gail
-gail_roc <- td_roc_plot(test_data, modelname="Gail", prsname="Genomics plc", plotname="Gail",
+gail_roc <- td_roc_plot(test_data, modelname="Gail", prsname="PRS_BC", plotname="Gail",
          varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk"))
 
+# Figure 2
+tiff(file.path(config$outputs$figures$final,"Figure2.tiff"), units="in", width=10, height=5, res=300)
 gridExtra::grid.arrange(TC_roc, gail_roc,
-                        ncol=2,
-                        top=paste0("Time-dependent ROC curves at 10 years follow up in test data (N=", nrow(test_data), ")"))
+                        ncol=2)
+dev.off()
 
@@
 dca_tc
-ggsave(file.path(config$outputs$figures, "TC_DCA.png"), 
+ggsave(file.path(config$outputs$figures$interim, "TC_DCA.png"), 
        scale=0.7, width=25, height=15, units="cm")
@@
 dca_gail
-ggsave(file.path(config$outputs$figures, "gail_DCA.png"), 
+ggsave(file.path(config$outputs$figures$interim, "gail_DCA.png"), 
        scale=0.7, width=25, height=15, units="cm")
 
-gridExtra::grid.arrange(dca_tc, dca_gail,
-                        ncol=2,
-                        top=paste0("Decision curves for 10-year risk from models with and without PRS_BC in test data (N=",
-                                   nrow(test_data), ")"))
+# Figure 3
+tiff(file.path(config$outputs$figures$final,"Figure3.tiff"), units="in", width=10, height=5, res=300)
+gridExtra::grid.arrange(dca_tc, dca_gail, ncol=2)
+dev.off()
 
@@
 ia_tc
-ggsave(file.path(config$outputs$figures, "TC_NIA.png"), 
+ggsave(file.path(config$outputs$figures$interim, "TC_NIA.png"), 
        scale=0.7, width=25, height=15, units="cm")
@@
 ia_gail
-ggsave(file.path(config$outputs$figures, "gail_NIA.png"), 
+ggsave(file.path(config$outputs$figures$interim, "gail_NIA.png"), 
        scale=0.7, width=25, height=15, units="cm")
 
-gridExtra::grid.arrange(ia_tc, ia_gail,
-                        ncol=2,
-                        top=paste0("Net interventions avoided per 100 women for 10-year risk from models with and without PRS_BC in test data (N=",
-                                   nrow(test_data), ")"))
+gridExtra::grid.arrange(ia_tc, ia_gail, ncol=2)
 
@@
 
-facet_data <- test_data %>% select(TC_10risk_cal, TC_prs_10risk, TEU_BrCa_status) %>% 
+# Tyrer-Cuzick
+facet_tc <- test_data %>% select(TC_10risk_cal, TC_prs_10risk, TEU_BrCa_status) %>% 
   pivot_longer(cols=c(TC_10risk_cal, TC_prs_10risk))
 
-density_data <- facet_data %>%
+density_data_tc <- facet_tc %>%
   mutate(
@@
 
-ggplot(density_data, 
+density_TC <- ggplot(density_data_tc, 
        aes(x=value, fill=Model)) + 
   geom_density(alpha=0.4, position="identity") +
-  ggtitle("Density plot of 10yr predicted risk by model in test data, split by cases and controls") +
   theme(legend.position = c(0.8, 0.8)) +
@@
 
-ggsave(file.path(config$outputs$figures, "TC_density_10yrrisk.png"), 
-       scale=0.7, width=25, height=15, units="cm")
-
-```
-
-```{r}
-
-facet_data <- test_data %>% select(gail_10risk_cal, gail_prs_10risk, TEU_BrCa_status) %>% 
+# Gail
+facet_gail <- test_data %>% select(gail_10risk_cal, gail_prs_10risk, TEU_BrCa_status) %>% 
   pivot_longer(cols=c(gail_10risk_cal, gail_prs_10risk))
 
-density_data <- facet_data %>%
+density_data_gail <- facet_gail %>%
   mutate(
@@
 
-ggplot(density_data, 
+density_gail <- ggplot(density_data_gail, 
        aes(x=value, fill=Model)) + 
   geom_density(alpha=0.4, position="identity") +
-  ggtitle("Density plot of 10yr predicted risk by model in test data, split by cases and controls") +
   theme(legend.position = c(0.8, 0.8)) +
@@
 
-ggsave(file.path(config$outputs$figures, "Gail_density_10yrrisk.png"), 
-       scale=0.7, width=25, height=15, units="cm")
+# Supplementary Figure 5
+tiff(file.path(config$outputs$figures$final,"SuppFigureS5.tiff"), units="in", width=8, height=10, res=300)
+gridExtra::grid.arrange(density_TC, density_gail, nrow=2)
+dev.off()
 

### 2024-02-13 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Label figures

@@
                         gail_prs_cal$plot + theme(legend.position="none"),
-                        ncol=3, widths=c(3.5, 3.5, 1))
+                        ncol=3, widths=c(3.5, 3.5, 1),
+                        top = grid::grid.text("Figure 1", x = 0, hjust = 0))
 dev.off()
@@
 gridExtra::grid.arrange(TC_roc, gail_roc,
-                        ncol=2)
+                        ncol=2,
+                        top = grid::grid.text("Figure 2", x = 0, hjust = 0))
 dev.off()
@@
 tiff(file.path(config$outputs$figures$final,"Figure3.tiff"), units="in", width=10, height=5, res=300)
-gridExtra::grid.arrange(dca_tc, dca_gail, ncol=2)
+gridExtra::grid.arrange(dca_tc, dca_gail, ncol=2,
+                        top = grid::grid.text("Figure 3", x = 0, hjust = 0))
 dev.off()
